---
title: "AICandRegressions"
author: "Catherine Otero"
date: "4/22/2022"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

The goal of this file is to run AICs to select variables to then incorporate into a regression evaluating AIM data and dams in Washington State.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries}
#check working directory
getwd()

#load packages
library(sf)
library(tidyverse)
library(nhdplusTools)
library(dplyr)
library(readr)
library(tmap)

# Set your ggplot theme
theme_set(theme_light())

```

Loading Scraped Data - may not need all of these
```{r load data}
WAAIM_tidy_dams <- st_read("./New Data/WAAIM_tidy_dams.csv")

catchment_data <- read.csv("./New Data/Dam_catchment_data.csv")

```

# Wrangle data
```{r wrangle data}
#check class of columns, we want numeric for analysis
sapply(WAAIM_tidy_dams, class)

#Specify columns want to change to numeric for WAAIM data
i <- c(5:58)  

#convert the columns indicated in i above to numeric
WAAIM_tidy_dams[ , i] <- apply(WAAIM_tidy_dams[ , i], 2,            
                    function(x) as.numeric(as.character(x)))

#check that all columns we want are numeric
sapply(WAAIM_tidy_dams, class)

#_________________________________________________________

#check class of catchment_data variables
sapply(catchment_data, class)

#Specify columns want to change to numeric for WAAIM data
j <- c(1,3)  

#convert the columns indicated in i above to numeric
catchment_data[ , j] <- apply(catchment_data[ , j], 2,            
                    function(x) as.numeric(as.character(x)))

#check class of catchment_data variables
sapply(catchment_data, class)

#rename Site_CD to SITE_CD so can join
names(catchment_data)[names(catchment_data)=="Site_CD"] <- "SITE_CD"
names(catchment_data)

#merge the dfs, must be the same class to merge
dam_data_merge = WAAIM_tidy_dams %>% left_join(catchment_data,by="SITE_CD")

#check which columns have NAs, only works on numeric columns
colSums(is.na(dam_data_clean))

#calculate a channel geometry column and add it onto df
#select columns of interest, gets rid of geometry so can save as a csv file
#drop rows for sites w/ NA for TotalP&N, invert variables, and D50
dam_data_clean <- dam_data_merge %>% 
  mutate(WDRatio = round(BNKFLL_WT / BNKFLL_HT)) %>% 
  select(OBJECTID, ReachLength, StreamOrder, VEG_CMPLXTY, RPRN_VEG_CNPY_CVR, RPRN_VEG_UNDRSTRY_CVR, RPRN_VEG_GC, NON_NTVE_WDY, NTVE_WDY, NON_NTVE_HRB, NTVE_HRB, SDGE_RSH, TotalN, TotalP, SpC, pH, INSTNT_TEMP, LWD_FREQ, LWDVolume, BNK_STBLTY, BNKFLL_HT, INCSN_HT, CHN_INCSN, BNKFLL_WT, WTTD_WT, WDRatio, OBSRVD_INVRT_RCHNSS, MACROINVRTBRTE_CNT, D50, Number_of_dams_upstream.x, Catchment_area_sqkm, Dams_per_sqkm) %>% 
  slice(-c(7,9,10,13,14,30,33,39,42,43,44))


#check data is complete, no NAs
colSums(is.na(dam_data_clean))

#save cleaned data as a csv
#write.csv(dam_data_clean, "./New Data/dam_data_clean.csv", row.names=FALSE)

```

# AIC Models and Regressions
We use AIC models to select which explanatory variables to use in a model. It can be tempting to throw all variables with available data into a model, but that can get confusing and not actually increase the explanatory power of a model. By using AIC to select variables, we can minimize covariates and end up with a simpler model that is informative and easier to interpret.

We perform regressions to see if and to what extent explanatory variables influence our dependent variables. We use the common p-value of below 0.05 to indicate significant relationships.
```{r AIC}
#read in data for the analysis
dam_analysis <- read.csv("./New Data/dam_data_clean.csv")

#AIC for total nitrogen as our dependent variable (DV)
TNAIC <- lm(data = dam_analysis, TotalN ~ ReachLength + StreamOrder + VEG_CMPLXTY + RPRN_VEG_CNPY_CVR + RPRN_VEG_UNDRSTRY_CVR + RPRN_VEG_GC + NON_NTVE_WDY + NTVE_WDY + NON_NTVE_HRB + NTVE_HRB + SDGE_RSH + TotalP + SpC + pH + INSTNT_TEMP + LWD_FREQ + LWDVolume + BNK_STBLTY + BNKFLL_HT + INCSN_HT + CHN_INCSN + BNKFLL_WT + WTTD_WT + Number_of_dams_upstream.x + OBSRVD_INVRT_RCHNSS + MACROINVRTBRTE_CNT + D50 + WDRatio + Catchment_area_sqkm + Dams_per_sqkm)

#Choose a model by AIC in a Stepwise Algorithm
step(TNAIC)
TNmodel <- lm(data = dam_analysis, TotalN ~ ReachLength + StreamOrder + VEG_CMPLXTY + RPRN_VEG_CNPY_CVR + RPRN_VEG_UNDRSTRY_CVR + RPRN_VEG_GC + NON_NTVE_HRB + NTVE_HRB + SDGE_RSH + TotalP + SpC + pH + INSTNT_TEMP + LWD_FREQ + LWDVolume + BNK_STBLTY + INCSN_HT + CHN_INCSN + BNKFLL_WT + WTTD_WT + Number_of_dams_upstream.x + OBSRVD_INVRT_RCHNSS + MACROINVRTBRTE_CNT + D50 + WDRatio + Catchment_area_sqkm + Dams_per_sqkm)
summary(TNmodel)

#Significant Variables
#at 0.05: RPRN_VEG_CNPY_CVR, RPRN_VEG_GC, RPRN_VEG_GC, LWD_FREQ, INCSN_HT, CHN_INCSN, WTTD_WT, Dams_per_sqkm 
#at 0.01: OBSRVD_INVRT_RCHNSS, WDRatio

#__________________________________________________

#AIC for total phosphorous as our DV
TPAIC <- lm(data = dam_analysis, TotalP ~ ReachLength + StreamOrder + VEG_CMPLXTY + RPRN_VEG_CNPY_CVR + RPRN_VEG_UNDRSTRY_CVR + RPRN_VEG_GC + NON_NTVE_WDY + NTVE_WDY + NON_NTVE_HRB + NTVE_HRB + SDGE_RSH + TotalN + SpC + pH + INSTNT_TEMP + LWD_FREQ + LWDVolume + BNK_STBLTY + BNKFLL_HT + INCSN_HT + CHN_INCSN + BNKFLL_WT + WTTD_WT + Number_of_dams_upstream.x + OBSRVD_INVRT_RCHNSS + MACROINVRTBRTE_CNT + D50 + WDRatio + Catchment_area_sqkm + Dams_per_sqkm)

#Choose a model by AIC in a Stepwise Algorithm
step(TPAIC)
TPmodel <- lm(data = dam_analysis, TotalP ~ ReachLength + StreamOrder + RPRN_VEG_CNPY_CVR + RPRN_VEG_GC + NON_NTVE_HRB + SDGE_RSH + INSTNT_TEMP + LWD_FREQ + BNK_STBLTY + MACROINVRTBRTE_CNT + D50 + WDRatio)
summary(TPmodel)

#Significant Variables
#at 0.05: WDRatio
#at 0.01: ReachLength, INSTNT_TEMP, LWD_FREQ, AMCROINVRTBRTE_CNT
#at 0.001: StreamOrder, BNK_STBLTY, D50

#___________________________________________-

#AIC for Observed Invertebrate Richness
Invert_AIC <- lm(data = dam_analysis, OBSRVD_INVRT_RCHNSS ~ ReachLength + StreamOrder + VEG_CMPLXTY + RPRN_VEG_CNPY_CVR + RPRN_VEG_UNDRSTRY_CVR + RPRN_VEG_GC + NON_NTVE_WDY + NTVE_WDY + NON_NTVE_HRB + NTVE_HRB + SDGE_RSH + TotalP + SpC + pH + INSTNT_TEMP + LWD_FREQ + LWDVolume + BNK_STBLTY + BNKFLL_HT + INCSN_HT + CHN_INCSN + BNKFLL_WT + WTTD_WT + Number_of_dams_upstream.x + TotalN + MACROINVRTBRTE_CNT + D50 + WDRatio + Catchment_area_sqkm + Dams_per_sqkm)

#Choose a model by AIC in a Stepwise Algorithm
step(Invert_AIC)
Invert_model <- lm(data = dam_analysis, OBSRVD_INVRT_RCHNSS ~ ReachLength + StreamOrder + VEG_CMPLXTY + RPRN_VEG_CNPY_CVR + RPRN_VEG_UNDRSTRY_CVR + RPRN_VEG_GC + NON_NTVE_HRB + NTVE_HRB + SDGE_RSH + TotalP + SpC + INSTNT_TEMP + LWD_FREQ + LWDVolume + BNK_STBLTY + INCSN_HT + CHN_INCSN + BNKFLL_WT + WTTD_WT + Number_of_dams_upstream.x + TotalN + MACROINVRTBRTE_CNT + D50 + WDRatio + Catchment_area_sqkm + Dams_per_sqkm)
summary(Invert_model)

#Significant Variables
#at 0.05: StreamOrder, RPRN_VEG_CNPY_CVR, RPRN_VEG_UNDRSTRY_CVR, RPRN_VEG_GC, NTVE_HRB, Number_of_dams_upstream.x, D50, Dams_per_sqkm
#at 0.01: LWD_FREQ, BNK_STBLTY, INCSN_HT, CHN_INCSN, WTTD_WT, WDRatio, Catchment_are_sqkm
#at 0.001: Non_NTVE_HRB, TotalN

#___________________________________________-

#AIC for Bank Full Height
WD_AIC <- lm(data = dam_analysis, WDRatio ~ ReachLength + StreamOrder + VEG_CMPLXTY + RPRN_VEG_CNPY_CVR + RPRN_VEG_UNDRSTRY_CVR + RPRN_VEG_GC + NON_NTVE_WDY + NTVE_WDY + NON_NTVE_HRB + NTVE_HRB + SDGE_RSH + TotalP + SpC + pH + INSTNT_TEMP + LWD_FREQ + LWDVolume + BNK_STBLTY + TotalN + INCSN_HT + CHN_INCSN + BNKFLL_WT + WTTD_WT + Number_of_dams_upstream.x + OBSRVD_INVRT_RCHNSS + MACROINVRTBRTE_CNT + D50 + BNKFLL_HT + Catchment_area_sqkm + Dams_per_sqkm)

#Choose a model by AIC in a Stepwise Algorithm
step(WD_AIC)
WD_model <- lm(data = dam_analysis, WDRatio ~ ReachLength + StreamOrder + VEG_CMPLXTY + RPRN_VEG_CNPY_CVR + RPRN_VEG_UNDRSTRY_CVR + RPRN_VEG_GC + NON_NTVE_HRB + NTVE_HRB + SDGE_RSH + TotalP + pH + SpC + INSTNT_TEMP + LWD_FREQ + LWDVolume + BNK_STBLTY + TotalN + INCSN_HT + CHN_INCSN + BNKFLL_WT + WTTD_WT + Number_of_dams_upstream.x + OBSRVD_INVRT_RCHNSS + MACROINVRTBRTE_CNT + D50 + BNKFLL_HT + Catchment_area_sqkm + Dams_per_sqkm)
summary(WD_model)

#Significant Variables
#at 0.05: RPRN_VEG_CNPY_CVR, RPRN_VEG_GC, NTVE_HRB, SDGE_RSH, BNK_STBLTY, TotalN, INCSN_HT, CHN_INCSN, WTTD_WT, Number_of_dams_upstream.x, D50, Dams_per_sqkm
#at 0.01: NON_NTVE_HRB, LWD_FREQ, OBSRVD_INVRT_RCHNSS

#___________________________________________-

#AIC for D50 aka particle sizes
D50AIC <- lm(data = dam_analysis, D50 ~ ReachLength + StreamOrder + VEG_CMPLXTY + RPRN_VEG_CNPY_CVR + RPRN_VEG_UNDRSTRY_CVR + RPRN_VEG_GC + NON_NTVE_WDY + NTVE_WDY + NON_NTVE_HRB + NTVE_HRB + SDGE_RSH + TotalP + SpC + pH + INSTNT_TEMP + LWD_FREQ + LWDVolume + BNK_STBLTY + BNKFLL_HT + INCSN_HT + CHN_INCSN + BNKFLL_WT + WTTD_WT + Number_of_dams_upstream.x + OBSRVD_INVRT_RCHNSS + MACROINVRTBRTE_CNT + TotalN + WDRatio + Catchment_area_sqkm + Dams_per_sqkm)

#Choose a model by AIC in a Stepwise Algorithm
step(D50AIC)
D50model <- lm(data = dam_analysis, D50 ~ StreamOrder + VEG_CMPLXTY + RPRN_VEG_CNPY_CVR + RPRN_VEG_GC + NTVE_WDY + NON_NTVE_HRB + NTVE_HRB + SDGE_RSH + TotalP + SpC + pH + INSTNT_TEMP + LWD_FREQ + LWDVolume + BNK_STBLTY + INCSN_HT + CHN_INCSN + WTTD_WT + Number_of_dams_upstream.x + OBSRVD_INVRT_RCHNSS + MACROINVRTBRTE_CNT + TotalN + WDRatio + Catchment_area_sqkm + Dams_per_sqkm)
summary(D50model)

#Significant Variables
#at 0.05: StreamOrder, RPRN_VEG_CNPY_CVR, NON_NTVE_HRB, pH, INCSN_HT, OBSRVD_INVRT_RCHNSS, Dams_per_sqkm
#at 0.01: VEG_CMPLXTY, RPRN_VEG_GC, SDGE_RSH, LWD_FREQ, Number_of_dams_upstream.x, TotalN 
#at 0.001: WTTD_WT, WDRatio


#___________________________________________-

#AIC for Dams_per_sqkm
Dams_AIC <- lm(data = dam_analysis, Dams_per_sqkm ~ ReachLength + StreamOrder + VEG_CMPLXTY + RPRN_VEG_CNPY_CVR + RPRN_VEG_UNDRSTRY_CVR + RPRN_VEG_GC + NON_NTVE_WDY + NTVE_WDY + NON_NTVE_HRB + NTVE_HRB + SDGE_RSH + TotalP + SpC + pH + INSTNT_TEMP + LWD_FREQ + LWDVolume + BNK_STBLTY + BNKFLL_HT + INCSN_HT + CHN_INCSN + BNKFLL_WT + WTTD_WT + Number_of_dams_upstream.x + OBSRVD_INVRT_RCHNSS + MACROINVRTBRTE_CNT + TotalN + WDRatio + Catchment_area_sqkm + D50)

#Choose a model by AIC in a Stepwise Algorithm
step(Dams_AIC)
Dams_model <- lm(formula = Dams_per_sqkm ~ StreamOrder + VEG_CMPLXTY + RPRN_VEG_CNPY_CVR + 
    RPRN_VEG_UNDRSTRY_CVR + RPRN_VEG_GC + NON_NTVE_HRB + NTVE_HRB + 
    SDGE_RSH + TotalP + pH + INSTNT_TEMP + LWD_FREQ + LWDVolume + 
    BNK_STBLTY + INCSN_HT + CHN_INCSN + BNKFLL_WT + WTTD_WT + 
    Number_of_dams_upstream.x + OBSRVD_INVRT_RCHNSS + MACROINVRTBRTE_CNT + 
    TotalN + WDRatio + Catchment_area_sqkm + D50, data = dam_analysis)
summary(Dams_model)

#Significant Variables
#at 0.05: VEG_CMPLXTY,RPRN_VEG_CNPY_CVR, NON_NTVE_HRB, SDGE_RSH, pH, LWDVolume, INCSN_HT , OBSRVD_INVRT_RCHNSS, TotalN, D50 
#at 0.01:  RPRN_VEG_GC, LWD_FREQ, WTTD_WT, Number_of_dams_upstream.x, WDRatio
#at 0.001: 


```

