---
title: "DamVariablesExtracting"
author: "Jackie Van Der Hout"
date: "4/21/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

The goal of this Rmd file is to extract from each AIM sample site:
* flowline distances from sample sites to upstream dams
* number of dams per square mile per catchment 
* number of dams upstream of each sample site

The above listed variables will be used for later analysis 


```{r}
library(sf)
library(tidyverse)
library(nhdplusTools)
library(dplyr)
library(readr)
library(mapview)

```

Loading Scraped Data
```{r}
WAAIM_tidy_utm <- st_read("./New Data/WAAIM_tidy.shp") %>% 
  st_transform(crs = 26910)

counties_sf_utm<- st_read('./New Data/cb_2018_us_county_20m.shp') %>% 
  filter(STATEFP == 53) %>%  #Filter for just WA Counties
  st_transform(crs = 26910)

dam_complete_raw_utm <- st_read("./New Data/dam_complete_raw.shp") %>% 
  st_transform(crs = 26910)

flowlines_utm <- st_read("./New Data/flowlinesall.shp") %>% 
  st_transform(crs = 26910)

catchments_utm <- st_read("./New Data/catchmentsall_true.shp") %>% 
  st_transform(crs = 26910)

```


Number of dams upstream of each sample site 
```{r}
#selecting the counties that intersect with the catchments 

county_intersect <- counties_sf_utm[lengths(st_intersects(counties_sf_utm, WAAIM_tidy_utm)) > 0,]
dam_intersect <- dam_complete_raw_utm[lengths(st_intersects(dam_complete_raw_utm, county_intersect)) > 0,]

dam_points_buf <- st_buffer(dam_complete_raw_utm, 15)
dam_intersect <- st_intersection(catchments_utm, dam_points_buf)

ggplot()+
  geom_sf(dam_intersect)+
  geom_sf(data = counties_sf_utm)+
  geom_sf(data = catchments_utm)

mapview::mapview(catchments_utm)+mapview(dam_intersect)

#get the area of the catchment areas
#join the dams together to the catchment areas 
#divide the two columns 

```


Flowline distance extracting 
*A - Snapping dam points to flowlines
(this has not worked for me and is not worth spending much more time on IMO)
```{r}

#st_snap function might help 

#taking the inside of the below function and trying to do it piece by piece 
# buffer the points by the tolerance
dam_points_buf <- st_buffer(dam_intersect, 15)
  
# intersect the line with the buffer
line_intersect <- st_intersection(flowlines_utm, dam_points_buf)

#use ndhplus to find the nearest point on a line 

mapview::mapview(line_intersect)

# convert mutlinestrings (more than one road segment) into linestrings
#line_intersect <- do.call(rbind,lapply(1:nrow(line_intersect),function(x){st_cast(line_intersect[x,],"LINESTRING")}))

# for each line intersection, calculate the nearest point on that line to our gps point
nearest_pt <- do.call(rbind,lapply(seq_along(points$id), ifunc))
  
nearest_pt <- points[points$id==i,] %>%  
  st_nearest_points(line_intersect[line_intersect$id==i,]) %>% 
  st_sf %>%
  st_cast('POINT') %>% 
  mutate(id = i) 

ifunc <- function(i){
    points[points$id==i,] %>%  
    st_nearest_points(line_intersect[line_intersect$id==i,]) %>% 
    st_sf %>%
    st_cast('POINT') %>% 
    mutate(id = i)
}

points %>% View()

nearest_pt <- do.call(rbind,lapply(seq_along(points$id), function(i){
    points[points$id==i,] %>%  st_nearest_points(line_intersect[line_intersect$id==i,]) %>% st_sf %>%
      st_cast('POINT') %>% mutate(id = i)
    }))

  
#__________________________________ 

custom_snap <- function(flowlines_utm, dam_points_buf, tolerance) {
  # intersect the line with the buffer
  line_intersect <- st_intersection(line, points_buf)
  # convert mutlinestrings (more than one road segment) into linestrings
  line_intersect <- do.call(rbind,lapply(1:nrow(line_intersect),function(x){st_cast(line_intersect[x,],"LINESTRING")}))

  # for each line intersection, calculate the nearest point on that line to our gps point
##this is not working because you can't subset a function 
## https://www.statology.org/r-object-of-type-closure-is-not-subsettable/ 
  nearest_pt <- do.call(rbind,lapply(seq_along(points$id), function(i){
    points[points$id==i,] %>%  st_nearest_points(line_intersect[line_intersect$id==i,]) %>% st_sf %>%
      st_cast('POINT') %>% mutate(id = i)
    }))

  nearest_pt<- nearest_pt[seq(2, nrow(nearest_pt), by = 2),] %>%
    mutate(option = 1:nrow(.))

  # find an unambiguous reference point with only one snap option
  unambiguous_pt <- nearest_pt %>%
    group_by(id) %>%
    mutate(count = n()) %>%
    ungroup() %>%
    filter(count == 1) %>%
    slice(1)

  # calculate network distance along our line to each snapped point
  dists <- rgeos::gProject(as(line,'Spatial'), as(nearest_pt,'Spatial'))
  # join back to nearest points data
  dists <- nearest_pt %>% cbind(dists)

  # we want to recursively do the following:
  # 1. calculate the network distance from our unambiguous reference point to the next id point in the data
  # 2. keep the snapped point for that id that was closest *along the network*  to the previous id
  # 3. set the newly snapped point as our reference point
  # 4. repeat

  # get distances from our reference point to the next point id
  for(i in unambiguous_pt$id:(max(dists$id)-1)){
    next_dist <- which.min(abs(dists[dists$id== i +1,]$dists - dists[dists$id== unambiguous_pt$id,]$dists ))
    next_option <- dists[dists$id== i +1,][next_dist,]$option
    nearest_pt <- nearest_pt %>% filter(id != i+1 | option == next_option)
    unambiguous_pt <- nearest_pt %>% filter(id ==i+1 & option == next_option)
    dists <- nearest_pt %>% cbind(dists = rgeos::gProject(as(line,'Spatial'), as(nearest_pt,'Spatial')))
  }

  # and in the reverse direction
  for(i in unambiguous_pt$id:(min(dists$id)+1)){
    next_dist <- which.min(abs(dists[dists$id== i -1,]$dists - dists[dists$id== unambiguous_pt$id,]$dists ))
    next_option <- dists[dists$id== i -1,][next_dist,]$option
    nearest_pt <- nearest_pt %>% filter(id != i-1 | option == next_option)
    unambiguous_pt <- nearest_pt %>% filter(id ==i-1 & option == next_option)
    dists <- nearest_pt %>% cbind(dists = rgeos::gProject(as(line,'Spatial'), as(nearest_pt,'Spatial')))
  }

  # transform back into lat/lng
  snapped_points <- nearest_pt %>%
    st_transform(4326)

  return(snapped_points)
}

flowlines_snap <- custom_snap(flowlines_utm, dam_intersect, 15) %>%
  cbind(st_coordinates(.))
```

* B - attaching AIM sites to flowlines (repeat above)
```{r}

```

* C - subtracting distances on flowlines
```{r}

```

* D - creating a new dataframe with including distances from each dam for every AIM sample site 
```{r}

```

Functions and intersecting dams and catchments
```{r}
catchment_extract <- function(site_id, waaim_data){
  selectingsites <- filter(waaim_data, SITE_CD == site_id) #takes site ID and filters dataset 
  MIDLAT <- selectingsites$MIDLAT
  MIDLONG <- selectingsites$MIDLONG
  start_point <- st_sfc(st_point(c(MIDLONG, MIDLAT)), crs = 4269)
  start_comid <- discover_nhdplus_id(start_point)
  NLDI <- navigate_nldi(list(featureSource = "comid", featureID = start_comid), 
                          mode = "upstreamTributaries", 
                          distance_km = 1000)
  subset_file <- tempfile(fileext = ".gpkg")
  subset <- subset_nhdplus(comids = as.integer(NLDI$UT$nhdplus_comid), 
                         output_file = subset_file, #the above temporary file 
                         nhdplus_data = "download", 
                         flowline_only = FALSE, 
                         return_data = TRUE, overwrite = TRUE)
  catchment <- subset$CatchmentSP
  return(catchment)
}

Dam_catchment_intersect <- function(CATCHMENT, DAM_DATA){
  DAM_DATA_UTM <- st_transform(DAM_DATA, crs = 26910)
  CATCHMENT_UTM <- st_transform(CATCHMENT, crs = 26910)
  DAM_SUBSET <- DAM_DATA_UTM[lengths(
  st_intersects(DAM_DATA_UTM, CATCHMENT_UTM)) > 0,]
  return(DAM_SUBSET)
}

Dam_WAAIM_Join_function <- function(WAAIM_ROW_NUM, WAAIM_DATA, DAM_DATA){
  site_id <- WAAIM_DATA$SITE_CD[WAAIM_ROW_NUM]
  site_catchment <- catchment_extract(site_id, WAAIM_DATA)
  site_dams <- Dam_catchment_intersect(site_catchment, DAM_DATA)
  dams_number <- dim(site_dams)[1]
  return(dams_number)
}

result <- data.frame(matrix(nrow = 45, ncol = 4))
for(i in 1:dim(WAAIM_tidy_utm)[1] ){
result[i,1] <- WAAIM_tidy_utm$SITE_CD[i]
site_catchment <- catchment_extract(result[i,1],WAAIM_tidy_utm)
site_dams <- Dam_catchment_intersect(site_catchment, dam_complete_raw_utm)
result[i,2] <- dim(site_dams)[1]
result[i,3] <- sum(site_catchment$areasqkm)
result[i,4] <- ((result[i,2])/(result[i,3]))
}
result <- result %>% 
  rename("Site_CD" = X1,
         "Number_of_dams_upstream" = X2,
         "Catchment_area_sqkm" = X3,
         "Dams_per_sqkm" = X4)

write.csv(result, "./New Data/Dam_catchment_data.csv")
        
```

