---
title: "Washington Dams and AIM Gage Sites"
author: "Jackie Van Der Hout, Andrew Friedman-Herring, & Catherine Otero"
date: "4/11/2022"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(lubridate)
library(sf)
library(corrplot)
library(dplyr)
theme_set(theme_light())
```

## Research Questions

- How many dams are upstream of the gage sites?
- How do the variables measured at each gage site correlate?
- How do the number of dams correlate with the AIM variables?

## Why ask?

- See the extent of the damming in Washington
- See the impact of dams on AIM variables

## What data did we use?
- BLM Assessment, Inventory, and
Monitoring (AIM) Data
-
-

## Water Quality Correlation Plot
```{r correlation plots, echo=FALSE}
WAAIM_tidy_dams_merge <- read.csv("./New Data/WAAIM_tidy_dams_merge.csv")

water_quality <- WAAIM_tidy_dams_merge %>% 
  select(c(23:30, 46))

water_corr <- cor(water_quality)
corrplot(water_corr, type = "upper", diag = FALSE, tl.cex = 0.8, tl.col = "black")
```

## Biodiversity and Riparian Habitat Quality Correlation Plot
```{r}
#selecting Biodiversity and Riparian Habitat Quality columns and number of dams and run correlation
#left out non-native woody veg b/c 0 for all sites
riparian_quality <- WAAIM_tidy_dams_merge %>% 
  select(c(11:14, 16:22, 46))

rip_corr <- cor(riparian_quality)
corrplot(rip_corr, type = "upper", diag = FALSE, tl.cex = 0.8, tl.col = "black")
```

## Watershed Function and Instream Habitat Quality Correlation Plot
```{r}

#selecting Watershed Function and Instream Habitat Quality and number of dams and run correlation
instream_quality <- WAAIM_tidy_dams_merge %>% 
  select(c(31:46))

inst_corr <- cor(instream_quality)
corrplot(inst_corr, type = "upper", diag = FALSE, tl.cex = 0.8, tl.col = "black")
```


## Next Steps
- Running an AIC
- Running regression

