---
title: "Washington Dams and AIM Gage Sites"
author: "Jackie Van Der Hout, Andrew Friedman-Herring, & Catherine Otero"
date: "4/11/2022"
output: ioslides_presentation
---
More info on slides in R here: https://bookdown.org/yihui/rmarkdown/ioslides-presentation.html

```{r setup, include=FALSE}
knitr::opts_chunk$set(include = FALSE)
library(tidyverse)
library(lubridate)
library(sf)
library(corrplot)
library(dplyr)
theme_set(theme_light())
```

```{r data scraping, include=FALSE}
#uploading datasets 
AIMraw <- st_read("https://gis.blm.gov/arcgis/rest/services/hydrography/BLM_Natl_AIM_AquADat/MapServer/0/query?where=1%3D1&text=&objectIds=&time=&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=*&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson&__ncforminfo=VmJvrSyZd1Eo2ftYuiZ2Csv3FXy8H7XoE-xtqGWFdpLoXXM_QrLBbJY-BtmB02LSjkYezZWjgMjkUQTDa0n0d7ugIiEzsdKzv1tB192rp3qdK8SWS-T_Vwxqlc6whlbENi3rrg4YqK40tpTjt0dJVWg3xwroBXxhGKT92BiR7Culk8jsk8g8cxPipo1GHHZ4oF9OZUqX8NX8gOXR3zsQRZxWmOIrkcEnN_JahqrcNjjDJXXh9ysVhdPK7UcQiEvMZy-n3fIQzqq-3YFqLGCq4ffzvu0ww4Pie6NuZD2CDkf9rilS-omAyiQq1LFJ0wVAl6PywwAseMQ4f6cDJXwLDn3BcTI3WP22jrsBYXm8SzrUZs5SGhJtMOOHN1-UMdNDFrO9rqMUXOQCTyKsmhFUFerLorYSAmMVohzr7cD03LvvdSkAQXKQpHQIcFuRqac69tN3eVlyV2IfsAA8yEcpU1GJKJx4jhwaVdXX9Sv4P1ACPp9mj3yqQ5zJqgAg0aFS6Z2O4GEd2B970yfKnbeicnbkgJTSCx0BQy3gtgkxCKujAYcWeMkvV5-dYcNF7lYQRaYvX8Eod5GUIbo8N8VNPF_USB9ZCQnlesBXLVMHXSOwH_bSq49wabgeZ2ZT7aZbJ1X8M_lXzzFieFrPzSKoOK_8_Bqcfde-Hdo8Jupndsn17KsTbpSH51tbBXOC6-3NJ0oBQuZAPkZ1I3pzzqb8sTFxIpYo2ia5lSHrZ5H-gxd753QRaQJBk1xWTyWgy8mON5hZYexZrjjUH4_C_SbY0Q%3D%3D")

WAAIM <- AIMraw %>% 
  filter(PRJCT == "WA_SpokaneDO")

WAAIM_tidy <- WAAIM %>% 
  select(OBJECTID, SITE_CD, "StreamName" = STRM_NM, MIDLAT, MIDLONG, "StreamOrder" = STRM_ORDR, BTMLAT, BTMLONG, TRLAT, TRLONG, "ReachLength" = TOT_RCH_LEN, "PercentCover" = PCT_OVRHD_CVR, VEG_CMPLXTY, RPRN_VEG_CNPY_CVR, RPRN_VEG_UNDRSTRY_CVR, RPRN_VEG_GC, NON_NTVE_WDY, NTVE_WDY, NON_NTVE_HRB, NTVE_HRB, SDGE_RSH, OBSRVD_INVRT_RCHNSS, EXPCTD_INVRT_RCHNSS, MACROINVRTBRTE_CNT, "TotalN" = TTL_N, "PredictedN" = PRD_TTL_N, "TotalP" = TTL_P, "PredictedP" = PRD_TTL_P, "SpC" = SPCFC_CNDCTNCE, "PredictedSpC" = PRD_SPCFC_CNDCTNCE, pH, INSTNT_TEMP, "PercentPools" = PCT_PL, "AvgPoolDepth" = RES_PL_DEP, "NumberPools" = NM_PL, LWD_FREQ, "LWDVolume" = LWD_VL, "PercentFines2" = PCT_FN2, "PercentFines6" = PCT_FN6, D16, D84, D50, BNK_STBLTY, BNK_CVR_LWD, BNKFLL_HT, INCSN_HT, CHN_INCSN, THLWG_DEP_MN, PCT_DRY, BNKFLL_WT, WTTD_WT, "FloodWidth" = FLD_WT, "EntrenchmentRatio" = ENTRCH, SLPE, SNSTY, BVR_SGN, geometry)

rm(AIMraw)
rm(WAAIM)

#Import Dam Data
dam_raw_0 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/0/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_0 <-  mutate(dam_raw_0, "Fish.Passage.Catagory" = rep("Not_a_barrier"))

dam_raw_1 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/1/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_1 <- mutate(dam_raw_1, "Fish.Passage.Catagory" = rep("Partial_Fish_Passage_Blockage"))

dam_raw_2 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/2/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_2 <- mutate(dam_raw_2, "Fish.Passage.Catagory" = rep("Total_Fish_Passage_Blockage"))

dam_raw_3 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/3/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_3 <- mutate(dam_raw_3, "Fish.Passage.Catagory" = rep("Barrier_Unknown_Percent_Passable"))

dam_raw_4 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/4/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_4 <- mutate(dam_raw_4, "Fish.Passage.Catagory" = rep("Diversion"))

dam_raw_5 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/5/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_5 <- mutate(dam_raw_5, "Fish.Passage.Catagory" = rep("Natural_Barrier_Verified"))

dam_raw_6 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/6/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_6 <- mutate(dam_raw_6, "Fish.Passage.Catagory" = rep("Natural_Barrier_Not_Field_Verified"))

dam_raw_7 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/7/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_7 <- mutate(dam_raw_7, "Fish.Passage.Catagory" = rep("On_A_Non_Fish_Bearing_Stream"))

dam_raw_8 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/8/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_8 <- mutate(dam_raw_8, "Fish.Passage.Catagory" = rep("Unknown"))

dam_complete_raw <- rbind(dam_raw_0, dam_raw_1, dam_raw_2,
                          dam_raw_3, dam_raw_4, dam_raw_5,
                          dam_raw_6, dam_raw_7, dam_raw_8)

rm(dam_raw_0, dam_raw_1, dam_raw_2,
    dam_raw_3, dam_raw_4, dam_raw_5,
                          dam_raw_6, dam_raw_7, dam_raw_8)

#extracting flowlines and catchments

flowline_extract <- function(site_id){
  selectingsites <- filter(WAAIM_tidy, SITE_CD == site_id) #takes site ID and filters dataset 
  MIDLAT <- selectingsites$MIDLAT
  MIDLONG <- selectingsites$MIDLONG
  start_point <- st_sfc(st_point(c(MIDLONG, MIDLAT)), crs = 4269)
  start_comid <- discover_nhdplus_id(start_point)
  NLDI <- navigate_nldi(list(featureSource = "comid", featureID = start_comid), 
                          mode = "upstreamTributaries", 
                          distance_km = 1000)
  subset_file <- tempfile(fileext = ".gpkg")
  subset <- subset_nhdplus(comids = as.integer(NLDI$UT$nhdplus_comid), 
                         output_file = subset_file, #the above temporary file 
                         nhdplus_data = "download", 
                         flowline_only = FALSE, 
                         return_data = TRUE, overwrite = TRUE)
  flowline <- subset$NHDFlowline_Network
  #catchment <- subset$CatchmentSP
  #waterbody <- subset$NHDWaterbody
  #waterlist <- list(flowline, catchment, waterbody)
  #names(waterlist) = c("flowline", "catchment", "waterbody")
  return(flowline)
}

catchment_extract <- function(site_id){
  selectingsites <- filter(WAAIM_tidy, SITE_CD == site_id) #takes site ID and filters dataset 
  MIDLAT <- selectingsites$MIDLAT
  MIDLONG <- selectingsites$MIDLONG
  start_point <- st_sfc(st_point(c(MIDLONG, MIDLAT)), crs = 4269)
  start_comid <- discover_nhdplus_id(start_point)
  NLDI <- navigate_nldi(list(featureSource = "comid", featureID = start_comid), 
                          mode = "upstreamTributaries", 
                          distance_km = 1000)
  subset_file <- tempfile(fileext = ".gpkg")
  subset <- subset_nhdplus(comids = as.integer(NLDI$UT$nhdplus_comid), 
                         output_file = subset_file, #the above temporary file 
                         nhdplus_data = "download", 
                         flowline_only = FALSE, 
                         return_data = TRUE, overwrite = TRUE)
  #flowline <- subset$NHDFlowline_Network
  catchment <- subset$CatchmentSP
  #waterbody <- subset$NHDWaterbody
  #waterlist <- list(flowline, catchment, waterbody)
  #names(waterlist) = c("flowline", "catchment", "waterbody")
  return(catchment)
}

site_name_vector <- WAAIM_tidy %>% pull(SITE_CD)

flowline_outer_frame <- data.frame(NULL)
for(cur_site_name in site_name_vector){ 
  cur_function_result <- flowline_extract(cur_site_name)
  head(cur_function_result)
  flowline_outer_frame <- rbind(flowline_outer_frame,cur_function_result)
}

catchment_outer_frame <- data.frame(NULL)
for(cur_site_name in site_name_vector){ 
  cur_function_result <- flowline_extract(cur_site_name)
  head(cur_function_result)
  catchment_outer_frame <- rbind(catchment_outer_frame,cur_function_result)
}

```

# **Research Questions**

- How many dams are upstream of the gage sites?
- How do the variables measured at each gage site correlate?
- How do the number of dams correlate with the AIM variables?

## *Why ask?*

- See the extent of the damming in Washington
- See the impact of dams on AIM variables

# **What data did we use?**
- BLM Assessment, Inventory, and
Monitoring (AIM) Data
- Washington State Department of Ecology Inventory of Statewide Dams
- National Hydrography Dataset of Flowlines and Catchments, developed by the US EPA and USGS

# **Data Wrangling**
- Data scraping publicly available geospatial data for AIM and dam data
- Using NHDplustools package to extract hydro data from the NHD

# **Analysis of the Data**
- Analytical approaches 
- Correlations among and within variables
- Geospatial tools to assess number of dams in catchments flowing into sample sites 

# **Results of the Analysis**

## Where are the AIM sample sites and their catchments? 
```{r}
tmap_mode("view")
tm_shape(outer_frame) + tm_lines() +
  tm_shape(WAAIM_tidy) + tm_dots() 
  #tm_shape(dam_complete_raw)+ tm_dots(col="red")

```


## Where are the dams in Washington State? 
```{r}

```

## Catchments, Dams, and Sample Sites
```{r}

```


## Water Quality Correlation Plot
```{r correlation plots, echo=FALSE}
WAAIM_tidy_dams_merge <- read.csv("./New Data/WAAIM_tidy_dams_merge.csv")

water_quality <- WAAIM_tidy_dams_merge %>% 
  select(c(23:30, 46))

water_corr <- cor(water_quality)
corrplot(water_corr, type = "upper", diag = FALSE, tl.cex = 0.8, tl.col = "black")
```

## Biodiversity and Riparian Habitat Quality Correlation Plot
```{r}
#selecting Biodiversity and Riparian Habitat Quality columns and number of dams and run correlation
#left out non-native woody veg b/c 0 for all sites
riparian_quality <- WAAIM_tidy_dams_merge %>% 
  select(c(11:14, 16:22, 46))

rip_corr <- cor(riparian_quality)
corrplot(rip_corr, type = "upper", diag = FALSE, tl.cex = 0.8, tl.col = "black")
```

## Watershed Function and Instream Habitat Quality Correlation Plot
```{r}

#selecting Watershed Function and Instream Habitat Quality and number of dams and run correlation
instream_quality <- WAAIM_tidy_dams_merge %>% 
  select(c(31:46))

inst_corr <- cor(instream_quality)
corrplot(inst_corr, type = "upper", diag = FALSE, tl.cex = 0.8, tl.col = "black")
```

#

# Next Steps
- Running an AIC
- Running regression

