---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: "Insert title of project here"
subtitle: "Web address for GitHub repository"
author: "Name"
fontsize: 12pt
mainfont: Times New Roman

---

### notes of things to update: 
*replace data scraping below with spatial versions of AIM, dam, flow, and catchment data
*add figure captions with fig.cap 
*insert images of the data sources under data section
*add metadata links to data section
*break up analysis into seperate pages as makes sense
*add writing describing each of the sections
*fill out lessons learned and further research interests in conclusion 
*Write 1-2 paragraph(s) detailing the rationale for your study. This should include both the context of the topic as well as a rationale for your choice of dataset (reason for location, variables, etc.). You may choose to include citations if you like (optional).


\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
# Check working directory
getwd()
# Load your packages
library(tidyverse)
library(lubridate)
library(sf)
library(corrplot)
library(dplyr)
library(nhdplusTools)
library(tmap)
tmap_mode("view") #makes map plots interactive
library(dataRetrieval)
library(maps)

# Set your ggplot theme
theme_set(theme_light())
```



```{r AIM & Dam data scraping}
# Import AIM Data
AIMraw <- st_read("https://gis.blm.gov/arcgis/rest/services/hydrography/BLM_Natl_AIM_AquADat/MapServer/0/query?where=1%3D1&text=&objectIds=&time=&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=*&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson&__ncforminfo=VmJvrSyZd1Eo2ftYuiZ2Csv3FXy8H7XoE-xtqGWFdpLoXXM_QrLBbJY-BtmB02LSjkYezZWjgMjkUQTDa0n0d7ugIiEzsdKzv1tB192rp3qdK8SWS-T_Vwxqlc6whlbENi3rrg4YqK40tpTjt0dJVWg3xwroBXxhGKT92BiR7Culk8jsk8g8cxPipo1GHHZ4oF9OZUqX8NX8gOXR3zsQRZxWmOIrkcEnN_JahqrcNjjDJXXh9ysVhdPK7UcQiEvMZy-n3fIQzqq-3YFqLGCq4ffzvu0ww4Pie6NuZD2CDkf9rilS-omAyiQq1LFJ0wVAl6PywwAseMQ4f6cDJXwLDn3BcTI3WP22jrsBYXm8SzrUZs5SGhJtMOOHN1-UMdNDFrO9rqMUXOQCTyKsmhFUFerLorYSAmMVohzr7cD03LvvdSkAQXKQpHQIcFuRqac69tN3eVlyV2IfsAA8yEcpU1GJKJx4jhwaVdXX9Sv4P1ACPp9mj3yqQ5zJqgAg0aFS6Z2O4GEd2B970yfKnbeicnbkgJTSCx0BQy3gtgkxCKujAYcWeMkvV5-dYcNF7lYQRaYvX8Eod5GUIbo8N8VNPF_USB9ZCQnlesBXLVMHXSOwH_bSq49wabgeZ2ZT7aZbJ1X8M_lXzzFieFrPzSKoOK_8_Bqcfde-Hdo8Jupndsn17KsTbpSH51tbBXOC6-3NJ0oBQuZAPkZ1I3pzzqb8sTFxIpYo2ia5lSHrZ5H-gxd753QRaQJBk1xWTyWgy8mON5hZYexZrjjUH4_C_SbY0Q%3D%3D")

WAAIM <- AIMraw %>% 
  filter(PRJCT == "WA_SpokaneDO")

WAAIM_tidy <- WAAIM %>% 
  select(OBJECTID, SITE_CD, "StreamName" = STRM_NM, MIDLAT, MIDLONG, "StreamOrder" = STRM_ORDR, BTMLAT, BTMLONG, TRLAT, TRLONG, "ReachLength" = TOT_RCH_LEN, "PercentCover" = PCT_OVRHD_CVR, VEG_CMPLXTY, RPRN_VEG_CNPY_CVR, RPRN_VEG_UNDRSTRY_CVR, RPRN_VEG_GC, NON_NTVE_WDY, NTVE_WDY, NON_NTVE_HRB, NTVE_HRB, SDGE_RSH, OBSRVD_INVRT_RCHNSS, EXPCTD_INVRT_RCHNSS, MACROINVRTBRTE_CNT, "TotalN" = TTL_N, "PredictedN" = PRD_TTL_N, "TotalP" = TTL_P, "PredictedP" = PRD_TTL_P, "SpC" = SPCFC_CNDCTNCE, "PredictedSpC" = PRD_SPCFC_CNDCTNCE, pH, INSTNT_TEMP, "PercentPools" = PCT_PL, "AvgPoolDepth" = RES_PL_DEP, "NumberPools" = NM_PL, LWD_FREQ, "LWDVolume" = LWD_VL, "PercentFines2" = PCT_FN2, "PercentFines6" = PCT_FN6, D16, D84, D50, BNK_STBLTY, BNK_CVR_LWD, BNKFLL_HT, INCSN_HT, CHN_INCSN, THLWG_DEP_MN, PCT_DRY, BNKFLL_WT, WTTD_WT, "FloodWidth" = FLD_WT, "EntrenchmentRatio" = ENTRCH, SLPE, SNSTY, BVR_SGN, geometry)

rm(AIMraw)
rm(WAAIM)


#Import Dam Data
dam_raw_0 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/0/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_0 <-  mutate(dam_raw_0, "Fish.Passage.Catagory" = rep("Not_a_barrier"))

dam_raw_1 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/1/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_1 <- mutate(dam_raw_1, "Fish.Passage.Catagory" = rep("Partial_Fish_Passage_Blockage"))

dam_raw_2 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/2/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_2 <- mutate(dam_raw_2, "Fish.Passage.Catagory" = rep("Total_Fish_Passage_Blockage"))

dam_raw_3 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/3/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_3 <- mutate(dam_raw_3, "Fish.Passage.Catagory" = rep("Barrier_Unknown_Percent_Passable"))

dam_raw_4 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/4/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_4 <- mutate(dam_raw_4, "Fish.Passage.Catagory" = rep("Diversion"))

dam_raw_5 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/5/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_5 <- mutate(dam_raw_5, "Fish.Passage.Catagory" = rep("Natural_Barrier_Verified"))

dam_raw_6 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/6/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_6 <- mutate(dam_raw_6, "Fish.Passage.Catagory" = rep("Natural_Barrier_Not_Field_Verified"))

dam_raw_7 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/7/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_7 <- mutate(dam_raw_7, "Fish.Passage.Catagory" = rep("On_A_Non_Fish_Bearing_Stream"))

dam_raw_8 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/8/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_8 <- mutate(dam_raw_8, "Fish.Passage.Catagory" = rep("Unknown"))

dam_complete_raw <- rbind(dam_raw_0, dam_raw_1, dam_raw_2,
                          dam_raw_3, dam_raw_4, dam_raw_5,
                          dam_raw_6, dam_raw_7, dam_raw_8)
rm(dam_raw_0, dam_raw_1, dam_raw_2,
    dam_raw_3, dam_raw_4, dam_raw_5,
                          dam_raw_6, dam_raw_7, dam_raw_8)

#write.csv(dam_complete_raw, "./New Data/dam_complete_raw.csv") this is to save time on our local computers to not have to data scrape every time



```

# Rationale and Research Questions

### Rationale: 

## Question 1: How many dams are upstream of each sample site?

## Question 2: How do the AIM variables correlate with each other?

## Question 3: How do the number of dams upstream of sample sites correlate with the AIM variables?

## *Why ask?*

- See the extent of the damming in Washington
- See the impact of dams on AIM variables
- Understand preliminary correlations for further analysis 

\newpage

# Dataset Information

Provide information on how the dataset for this analysis were collected, the data contained in the dataset, and any important pieces of information that are relevant to your analyses. This section should contain much of same information as the metadata file for the dataset but formatted in a way that is more narrative.

Describe how you wrangled your dataset in a format similar to a methods section of a journal article.

Add a table that summarizes your data structure (variables, units, ranges and/or central tendencies, data source if multiple are used, etc.). This table can be made in markdown text or inserted as a `kable` function in an R chunk. If the latter, do not include the code used to generate your table.

## **What data did we use?**
- BLM Assessment, Inventory, and
Monitoring (AIM) Data
- Washington State Department of Ecology Inventory of Statewide Dams
- National Hydrography Dataset of Flowlines and Catchments, developed by the US EPA and USGS

\newpage

# **Data Wrangling**
- Data scraping publicly available geospatial data for AIM and dam data
- Using NHDplustools package to extract hydro data from the NHD

\newpage

# **Exploratory Data Analysis**

Insert exploratory visualizations of your dataset. This may include, but is not limited to, graphs illustrating the distributions of variables of interest and/or maps of the spatial context of your dataset. Format your R chunks so that graphs are displayed but code is not displayed. Accompany these graphs with text sections that describe the visualizations and provide context for further analyses.

Each figure should be accompanied by a caption, and each figure should be referenced within the text. 

Scope: think about what information someone might want to know about the dataset before analyzing it statistically. How might you visualize this information? 

- Correlations among and within variables
- Geospatial tools to assess number of dams in catchments flowing into sample sites 

\newpage

# **Analysis**

Insert visualizations and text describing your main analyses. Format your R chunks so that graphs are displayed but code and other output is not displayed. Instead, describe the results of any statistical tests in the main text (e.g., "Variable x was significantly different among y groups (ANOVA; df = 300, F = 5.55, p < 0.0001)"). Each paragraph, accompanied by one or more visualizations, should describe the major findings and how they relate to the question and hypotheses. Divide this section into subsections, one for each research question.

Each figure should be accompanied by a caption, and each figure should be referenced within the text

# Where are the AIM sample sites and their catchments? 
```{r echo=FALSE}
ggplot(counties_sf)+
  geom_sf()+
  geom_sf(data = catchments, color = "blue", alpha = 0.3)+
  geom_sf(data = WAAIM_tidy, color = "green")

```


# Where are all of dams in Washington State? 
```{r Dams in Washington State, echo=FALSE}
ggplot(counties_sf_utm)+
  geom_sf()+
  geom_sf(data = dam_complete_raw_utm, color = "dark red")



```

# Catchments, Dams, and Sample Sites
```{r echo=FALSE}
ggplot(counties_sf)+
  geom_sf()+
  geom_sf(data = catchments, color = "blue", alpha = 0.3)+
  geom_sf(data = Dam_complete_catchment, color = "red", size = 0.5)+
  geom_sf(data = WAAIM_tidy, color = "green")
```


# Water Quality Correlation Plot
```{r correlation plots, echo=FALSE}
WAAIM_tidy_dams_merge <- read.csv("../New Data/WAAIM_tidy_dams_merge.csv")

water_quality <- WAAIM_tidy_dams_merge %>% 
  select(c(23:30, 46))

water_corr <- cor(water_quality)
corrplot(water_corr, type = "upper", diag = FALSE, tl.cex = 0.8, tl.col = "black")
```

# Biodiversity and Riparian Habitat Quality Correlation Plot
```{r echo=FALSE}
#selecting Biodiversity and Riparian Habitat Quality columns and number of dams and run correlation
#left out non-native woody veg b/c 0 for all sites
riparian_quality <- WAAIM_tidy_dams_merge %>% 
  select(c(11:14, 16:22, 46))

rip_corr <- cor(riparian_quality)
corrplot(rip_corr, type = "upper", diag = FALSE, tl.cex = 0.8, tl.col = "black")
```

# Watershed Function and Instream Habitat Quality Correlation Plot
```{r echo=FALSE}

#selecting Watershed Function and Instream Habitat Quality and number of dams and run correlation
instream_quality <- WAAIM_tidy_dams_merge %>% 
  select(c(31:46))

inst_corr <- cor(instream_quality)
corrplot(inst_corr, type = "upper", diag = FALSE, tl.cex = 0.8, tl.col = "black")
```

# **Next Steps**
- Running an AIC
- Running regression
- Calculating flowline distances with NHDtoolsplus
- Comparing on AIM variables between sample site catchments on the basis of # of dams, dams / square mile, and flowline distance to dams




# Data Exploration of AIM Data
```{r data exploration}
#repeat for both datasets
unique(AIMraw$PRJCT)

WAAIM <- AIMraw %>% 
  filter(PRJCT == "WA_SpokaneDO") 

SiteVisits <- WAAIM %>% 
  group_by(SITE_CD) %>% 
  summarise(columnname = n()) 
#all sites were only sampled once 

rm(AIMraw)

class()
colnames()
summary()
```
# Data Tidying for AIM Data
```{r data tidying}
WAAIM_tidy <- WAAIM %>% 
  select(OBJECTID, SITE_CD, "StreamName" = STRM_NM, MIDLAT, MIDLONG, "StreamOrder" = STRM_ORDR, BTMLAT, BTMLONG, TRLAT, TRLONG, "ReachLength" = TOT_RCH_LEN, "PercentCover" = PCT_OVRHD_CVR, VEG_CMPLXTY, RPRN_VEG_CNPY_CVR, RPRN_VEG_UNDRSTRY_CVR, RPRN_VEG_GC, NON_NTVE_WDY, NTVE_WDY, NON_NTVE_HRB, NTVE_HRB, SDGE_RSH, OBSRVD_INVRT_RCHNSS, EXPCTD_INVRT_RCHNSS, MACROINVRTBRTE_CNT, "TotalN" = TTL_N, "PredictedN" = PRD_TTL_N, "TotalP" = TTL_P, "PredictedP" = PRD_TTL_P, "SpC" = SPCFC_CNDCTNCE, "PredictedSpC" = PRD_SPCFC_CNDCTNCE, pH, INSTNT_TEMP, "PercentPools" = PCT_PL, "AvgPoolDepth" = RES_PL_DEP, "NumberPools" = NM_PL, LWD_FREQ, "LWDVolume" = LWD_VL, "PercentFines2" = PCT_FN2, "PercentFines6" = PCT_FN6, D16, D84, D50, BNK_STBLTY, BNK_CVR_LWD, BNKFLL_HT, INCSN_HT, CHN_INCSN, THLWG_DEP_MN, PCT_DRY, BNKFLL_WT, WTTD_WT, "FloodWidth" = FLD_WT, "EntrenchmentRatio" = ENTRCH, SLPE, SNSTY, BVR_SGN, geometry)

#remove WAAIM raw
rm(WAAIM)
```

#Creating Functions for extracting Catchment and flowlines from NHDSplus Tools
```{r}
flowline_extract <- function(site_id){
  selectingsites <- filter(WAAIM_tidy, SITE_CD == site_id) #takes site ID and filters dataset 
  MIDLAT <- selectingsites$MIDLAT
  MIDLONG <- selectingsites$MIDLONG
  start_point <- st_sfc(st_point(c(MIDLONG, MIDLAT)), crs = 4269)
  start_comid <- discover_nhdplus_id(start_point)
  NLDI <- navigate_nldi(list(featureSource = "comid", featureID = start_comid), 
                          mode = "upstreamTributaries", 
                          distance_km = 1000)
  subset_file <- tempfile(fileext = ".gpkg")
  subset <- subset_nhdplus(comids = as.integer(NLDI$UT$nhdplus_comid), 
                         output_file = subset_file, #the above temporary file 
                         nhdplus_data = "download", 
                         flowline_only = FALSE, 
                         return_data = TRUE, overwrite = TRUE)
  flowline <- subset$NHDFlowline_Network
  return(flowline)
}

catchment_extract <- function(site_id){
  selectingsites <- filter(WAAIM_tidy, SITE_CD == site_id) #takes site ID and filters dataset 
  MIDLAT <- selectingsites$MIDLAT
  MIDLONG <- selectingsites$MIDLONG
  start_point <- st_sfc(st_point(c(MIDLONG, MIDLAT)), crs = 4269)
  start_comid <- discover_nhdplus_id(start_point)
  NLDI <- navigate_nldi(list(featureSource = "comid", featureID = start_comid), 
                          mode = "upstreamTributaries", 
                          distance_km = 1000)
  subset_file <- tempfile(fileext = ".gpkg")
  subset <- subset_nhdplus(comids = as.integer(NLDI$UT$nhdplus_comid), 
                         output_file = subset_file, #the above temporary file 
                         nhdplus_data = "download", 
                         flowline_only = FALSE, 
                         return_data = TRUE, overwrite = TRUE)
  catchment <- subset$CatchmentSP
  return(catchment)
}
```

#Intersecting Dam Data with flowline + catchment data above AIM sites

```{r}
Dam_catchment_intersect <- function(CATCHMENT, DAM_DATA){
  DAM_DATA_UTM <- st_transform(DAM_DATA, crs = 26910)
  CATCHMENT_UTM <- st_transform(CATCHMENT, crs = 26910)
  DAM_SUBSET <- DAM_DATA_UTM[lengths(
  st_intersects(DAM_DATA_UTM, CATCHMENT_UTM)) > 0,]
  return(DAM_SUBSET)
}

Dam_WAAIM_Join_function <- function(WAAIM_ROW_NUM, WAAIM_DATA, DAM_DATA){
  site_id <- WAAIM_DATA$SITE_CD[WAAIM_ROW_NUM]
  site_catchment <- catchment_extract(site_id)
  site_dams <- Dam_catchment_intersect(site_catchment, DAM_DATA)
  dams_number <- dim(site_dams)[1]
  return(dams_number)
}

result <- data.frame(matrix(nrow = 45, ncol = 2))
for(i in 1:dim(WAAIM_tidy)[1] ){
result[i,1] <-  Dam_WAAIM_Join_function(i,WAAIM_tidy,dam_complete_raw)
result[i,2] <- WAAIM_tidy$SITE_CD[i]
}
result$SITE_CD <-  result$X2
result <- result %>% 
  rename("Number_of_dams_upstream" = X1)

WAAIM_tidy_dams <- left_join(WAAIM_tidy, result)

write.csv(WAAIM_tidy_dams, "./New Data/WAAIM_tidy_dams.csv")

#dams_upstream <- st_read("./New Data/WAAIM_tidy_dams.csv", options=c("X_POSSIBLE_NAMES=x","Y_POSSIBLE_NAMES=y"))
```

# Correlation Check on AIM Data

We run the correlation check to see what variables in the AIM data set are correlated with each other and also which variables are correlated with the number of dams upstream from the AIM gage site. Correlation plots can be used as a visualization to help choose what variables to include in regression models.
```{r correlation check}
#Specify columns want to change to numeric
#Change integer columns to numeric b/c corr can only be run on numeric columns

i <- c(1, 6, 17:22, 24, 35, 40:44)  

#convert the columns indicated in i above to numeric
WAAIM_tidy[ , i] <- apply(WAAIM_tidy[ , i], 2,            # Specify own function within apply
                    function(x) as.numeric(as.character(x)))

#check that all columns we want are integers
sapply(WAAIM_tidy, class)

#drop nas so can run a corr, narrows down our data
WAAIM_tidy <- drop_na(WAAIM_tidy)

#select for numeric columns to run the corr on
WAAIM_tidy_numeric <- WAAIM_tidy %>%
  select(c(1,4:47))

#drop the geometry column b/c cor only runs on numeric columns
WAAIM_tidy_numeric <- st_set_geometry(WAAIM_tidy_numeric, NULL)

#check if geometry column removed
head(WAAIM_tidy_numeric)

#read in updated data with number of upstream dams attached. Warning is okay.
WAAIM_tidy_dams <- st_read("./New Data/WAAIM_tidy_dams.csv")

head(WAAIM_tidy_dams)

#select out the number of dams upstream column
upstream_dams <- WAAIM_tidy_dams %>% 
  select(OBJECTID, Number_of_dams_upstream)

view(upstream_dams)

#change both columns to numeric

numb <- c(1,2)  

upstream_dams[ , numb] <- apply(upstream_dams[ , numb], 2,            # Specify own function within apply
                    function(x) as.numeric(as.character(x)))

sapply(upstream_dams, class)

#merge the dfs, must be the same class to merge
WAAIM_tidy_dams_merge = WAAIM_tidy_numeric %>% left_join(upstream_dams,by="OBJECTID")
WAAIM_tidy_dams_merge

head(WAAIM_tidy_dams_merge)

# create a correlation matrix. Necessitates only complete cases. Needs to be its own data matrix. Use corr to create the matrix so all variables are the columns and rows and their correlation between each other is calculated. The dark blue diagonal down the middle is b/c when something correlated against itself is 1. Used to see how covariates are related aka correlated
WAAIMcorr <- cor(WAAIM_tidy_dams_merge)

# plot correlation matrix.
corrplot(WAAIMcorr)

# customize so that only the upper half shows since the upper and lower (compared to the diagonal mirrors itself) with the type, diag = FALSE gets the self-correlation removed, endings change the text 
corrplot(WAAIMcorr, type = "upper", diag = FALSE, tl.cex = 0.8, tl.col = "black")

#Checking correlation between variables that are categorized together in the AIM metadata.

#selecting water quality columns and number of dams and run correlation
water_quality <- WAAIM_tidy_dams_merge %>% 
  select(c(23:30, 46))

water_corr <- cor(water_quality)
corrplot(water_corr)
corrplot(water_corr, type = "upper", diag = FALSE, tl.cex = 0.8, tl.col = "black")

#selecting Biodiversity and Riparian Habitat Quality columns and number of dams and run correlation
#left out non-native woody veg b/c 0 for all sites
riparian_quality <- WAAIM_tidy_dams_merge %>% 
  select(c(11:14, 16:22, 46))

rip_corr <- cor(riparian_quality)
corrplot(rip_corr)
corrplot(rip_corr, type = "upper", diag = FALSE, tl.cex = 0.8, tl.col = "black")

#selecting Watershed Function and Instream Habitat Quality and number of dams and run correlation
instream_quality <- WAAIM_tidy_dams_merge %>% 
  select(c(31:46))

inst_corr <- cor(instream_quality)
corrplot(inst_corr)
corrplot(inst_corr, type = "upper", diag = FALSE, tl.cex = 0.8, tl.col = "black")
```
# Results of Correlation Plots
In the water quality plot, we saw number of dams upstream correlate positively with total nitrogen and pH. In the biodiversity and riparian habitat quality plot, we saw number of dams upstream correlate positively with non-native herbaceous vegetation. In the watershed function and in-stream habitat quality plot we saw number of dams upstream correlate negatively with percent pools in stream and average bank cover. 

# Summary and Conclusions

Washington's rivers are saturated with dams. Since the AIM sites are one-off sampling occurrences, the sites provide a wealth of information but not a clear way to examine or compare data. 

Summarize your major findings from your analyses in a few paragraphs. What conclusions do you draw from your findings? Relate your findings back to the original research questions and rationale.

# Next Steps
The initial analyses we ran prepared us to further investigate the data. We plan to:
- Run an AIC to choose the most influential variables
- Run regression with the AIC determined variables
- Run a regression that includes stream order, dam type, number of dams upstream, and distance to dam with selected AIM variable as the dependent variable
- Calculate flowline distances with NHDtoolsplus
- Compare on AIM variables between sample site catchments on the basis of # of dams, dams / square mile, and flowline distance to dams

\newpage

# References
<add references here if relevant, otherwise delete this section> 
