---
title: "Analyzing the relationships between variables monitored at AIM sites in WA and the number of dams upstream"
author: "Catherine Otero, Jackie Van der Hout, Andrew Friedman-Herring"
output:
  html_document:
    df_print: paged
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
subtitle: Web address for GitHub repository
geometry: margin=2.54cm
fontsize: 12pt
mainfont: Times New Roman
---


```{r setup, include=FALSE}
# Check working directory
getwd()
# Load your packages
library(tidyverse)
library(lubridate)
library(sf)
library(corrplot)
library(dplyr)
library(nhdplusTools)
library(tmap)
tmap_mode("view") #makes map plots interactive
library(dataRetrieval)
library(maps)

# Set your ggplot theme
theme_set(theme_light())
```


# Rationale and Research Questions

### Rationale: 

We chose this project and these research questions specifically because water quality is an incredibly important issue, and we wanted to see how physical and biological water quality indicators collected by the BLM correlate with each other. Dams also have a very large impact on streams and rivers in the US, as they increase stream temperatures and provide conditions for algal blooms which can reduce water quality. Dams are also a very overlooked piece of infrastructure. By collecting the number of dams upstream of each monitoring site, we are able to add that value to the AIM data set and incorporate it with our analysis of how water quality variables correlate.

## Question 1: How many dams are upstream of each BLM Assessment, Inventory, and Monitoring (AIM) sample site in Washington State?

## Question 2: How do the AIM variables correlate with each other across all of the sites?

## Question 3: How do the number of dams upstream of sample sites correlate with the AIM variables across all of the sites?



 


\newpage

# Dataset Information

## **What data did we use?**

For our analysis, we used the BLM Assessment, Inventory, and Monitoring (AIM) Data, which we found and data scraped from online in a GeoJSON format. 
*add information about variables in data set*

We also used the Washington State Department of Ecology Inventory of Statewide Dams, which was also found and data scraped from online in a GeoJSON format. This information included latitudes and longitudes for all 57,000 dams in Washington State, as well as their fish barrier classification, which ranged from "Not a barrier" to "Complete Fish Barrier". 

Finally, we used the National Hydrography Dataset of Flowlines and Catchments developed by the US EPA and USGS to connect first two data set spatially. This dataset was used in the form of the NHDplusTools R package to find the catchment and flowline data upstream of each monitoring site. 




\newpage

# **Data Wrangling**

  The data wrangling phase of our project was the most time and code intensive, as we were trying to intersect two very large data sets spatially to create a new variable to add to the AIM data set. Because there were 45 AIM sites in Washington, we had to repeat this intersection for all 45 sites. 
  We accomplished this by writing a custom function that extracted the catchment above a monitoring site and saved it as a a spatial polygon object. We then wrote a 'for' loop that would apply this catchment extraction function for all 45 sites and save them as one combined spatial dataframe. We repeated this process for the flowlines above each monitoring site as well to be used in a future analysis. 
  We then wrote another set of functions that takes the catchment data, intersects it with the inventory of dams and returns the number of dams contained in that catchment. We then wrote another 'for' loop to apply this function to all 45 AIM sites, and return a data frame containing the number of dams above each AIM site. This was then joined to the original AIM dataframe, effectively adding a column to the AIM data set counting how many dams were above each monitoring site. 




\newpage


#Data Scraping
```{r AIM & Dam data scraping and saving as csv files, include=FALSE}
# Import AIM Data
AIMraw <- st_read("https://gis.blm.gov/arcgis/rest/services/hydrography/BLM_Natl_AIM_AquADat/MapServer/0/query?where=1%3D1&text=&objectIds=&time=&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=*&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson&__ncforminfo=VmJvrSyZd1Eo2ftYuiZ2Csv3FXy8H7XoE-xtqGWFdpLoXXM_QrLBbJY-BtmB02LSjkYezZWjgMjkUQTDa0n0d7ugIiEzsdKzv1tB192rp3qdK8SWS-T_Vwxqlc6whlbENi3rrg4YqK40tpTjt0dJVWg3xwroBXxhGKT92BiR7Culk8jsk8g8cxPipo1GHHZ4oF9OZUqX8NX8gOXR3zsQRZxWmOIrkcEnN_JahqrcNjjDJXXh9ysVhdPK7UcQiEvMZy-n3fIQzqq-3YFqLGCq4ffzvu0ww4Pie6NuZD2CDkf9rilS-omAyiQq1LFJ0wVAl6PywwAseMQ4f6cDJXwLDn3BcTI3WP22jrsBYXm8SzrUZs5SGhJtMOOHN1-UMdNDFrO9rqMUXOQCTyKsmhFUFerLorYSAmMVohzr7cD03LvvdSkAQXKQpHQIcFuRqac69tN3eVlyV2IfsAA8yEcpU1GJKJx4jhwaVdXX9Sv4P1ACPp9mj3yqQ5zJqgAg0aFS6Z2O4GEd2B970yfKnbeicnbkgJTSCx0BQy3gtgkxCKujAYcWeMkvV5-dYcNF7lYQRaYvX8Eod5GUIbo8N8VNPF_USB9ZCQnlesBXLVMHXSOwH_bSq49wabgeZ2ZT7aZbJ1X8M_lXzzFieFrPzSKoOK_8_Bqcfde-Hdo8Jupndsn17KsTbpSH51tbBXOC6-3NJ0oBQuZAPkZ1I3pzzqb8sTFxIpYo2ia5lSHrZ5H-gxd753QRaQJBk1xWTyWgy8mON5hZYexZrjjUH4_C_SbY0Q%3D%3D")

WAAIM <- AIMraw %>% 
  filter(PRJCT == "WA_SpokaneDO")

WAAIM_tidy <- WAAIM %>% 
  select(OBJECTID, SITE_CD, "StreamName" = STRM_NM, MIDLAT, MIDLONG, "StreamOrder" = STRM_ORDR, BTMLAT, BTMLONG, TRLAT, TRLONG, "ReachLength" = TOT_RCH_LEN, "PercentCover" = PCT_OVRHD_CVR, VEG_CMPLXTY, RPRN_VEG_CNPY_CVR, RPRN_VEG_UNDRSTRY_CVR, RPRN_VEG_GC, NON_NTVE_WDY, NTVE_WDY, NON_NTVE_HRB, NTVE_HRB, SDGE_RSH, OBSRVD_INVRT_RCHNSS, EXPCTD_INVRT_RCHNSS, MACROINVRTBRTE_CNT, "TotalN" = TTL_N, "PredictedN" = PRD_TTL_N, "TotalP" = TTL_P, "PredictedP" = PRD_TTL_P, "SpC" = SPCFC_CNDCTNCE, "PredictedSpC" = PRD_SPCFC_CNDCTNCE, pH, INSTNT_TEMP, "PercentPools" = PCT_PL, "AvgPoolDepth" = RES_PL_DEP, "NumberPools" = NM_PL, LWD_FREQ, "LWDVolume" = LWD_VL, "PercentFines2" = PCT_FN2, "PercentFines6" = PCT_FN6, D16, D84, D50, BNK_STBLTY, BNK_CVR_LWD, BNKFLL_HT, INCSN_HT, CHN_INCSN, THLWG_DEP_MN, PCT_DRY, BNKFLL_WT, WTTD_WT, "FloodWidth" = FLD_WT, "EntrenchmentRatio" = ENTRCH, SLPE, SNSTY, BVR_SGN, geometry)

rm(AIMraw)
rm(WAAIM)

WAAIM_tidy3 <- st_read("../New Data/WAAIM_tidy.csv", geometry_column = geometry)


st_write(WAAIM_tidy, "../New Data/WAAIM_tidy2.shp")
WAAIM_tidy2 <- st_read("../New Data/WAAIM_tidy2.shp")


#do the same with dam data seems to work
#however then the for loops with the flowline and catchment functions don't seem to work with these files? 
#also take a look at the CSVs for the flowlines and the catchments 

##write.csv(WAAIM_tidy, "../New Data/WAAIM_tidy.csv")

#Import Dam Data
dam_raw_0 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/0/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_0 <-  mutate(dam_raw_0, "Fish.Passage.Catagory" = rep("Not_a_barrier"))

dam_raw_1 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/1/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_1 <- mutate(dam_raw_1, "Fish.Passage.Catagory" = rep("Partial_Fish_Passage_Blockage"))

dam_raw_2 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/2/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_2 <- mutate(dam_raw_2, "Fish.Passage.Catagory" = rep("Total_Fish_Passage_Blockage"))

dam_raw_3 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/3/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_3 <- mutate(dam_raw_3, "Fish.Passage.Catagory" = rep("Barrier_Unknown_Percent_Passable"))

dam_raw_4 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/4/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_4 <- mutate(dam_raw_4, "Fish.Passage.Catagory" = rep("Diversion"))

dam_raw_5 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/5/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_5 <- mutate(dam_raw_5, "Fish.Passage.Catagory" = rep("Natural_Barrier_Verified"))

dam_raw_6 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/6/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_6 <- mutate(dam_raw_6, "Fish.Passage.Catagory" = rep("Natural_Barrier_Not_Field_Verified"))

dam_raw_7 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/7/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_7 <- mutate(dam_raw_7, "Fish.Passage.Catagory" = rep("On_A_Non_Fish_Bearing_Stream"))

dam_raw_8 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/8/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_8 <- mutate(dam_raw_8, "Fish.Passage.Catagory" = rep("Unknown"))

dam_complete_raw <- 
  rbind(dam_raw_0, dam_raw_1, dam_raw_2,
        dam_raw_3, dam_raw_4, dam_raw_5,
        dam_raw_6, dam_raw_7, dam_raw_8)

rm(dam_raw_0, dam_raw_1, dam_raw_2,
   dam_raw_3, dam_raw_4, dam_raw_5,
   dam_raw_6, dam_raw_7, dam_raw_8)

counties_sf<- st_read('../New Data/cb_2018_us_county_20m.shp') %>% 
  filter(STATEFP == 53) #Filter for just WA Counties

counties_sf_utm <- st_transform(counties_sf, crs = 26910)
dam_complete_raw_utm <- st_transform(dam_complete_raw, crs = 26910)
WAAIM_tidy_utm <- st_transform(WAAIM_tidy, crs = 26910)

#st_write(dam_complete_raw, "./New Data/dam_complete_raw.shp") #this is to save time on our local computers to not have to data scrape every time

#dam_read_in <- st_read("../New Data/dam_complete_raw.shp")

```
#Creating Functions for extracting Catchment and flowlines from NHDSplus Tools
```{r message=FALSE, warning=FALSE}
flowline_extract <- function(site_id){
  selectingsites <- filter(WAAIM_tidy, SITE_CD == site_id) #takes site ID and filters dataset 
  MIDLAT <- selectingsites$MIDLAT
  MIDLONG <- selectingsites$MIDLONG
  start_point <- st_sfc(st_point(c(MIDLONG, MIDLAT)), crs = 4269)
  start_comid <- discover_nhdplus_id(start_point)
  NLDI <- navigate_nldi(list(featureSource = "comid", featureID = start_comid), 
                          mode = "upstreamTributaries", 
                          distance_km = 1000)
  subset_file <- tempfile(fileext = ".gpkg")
  subset <- subset_nhdplus(comids = as.integer(NLDI$UT$nhdplus_comid), 
                         output_file = subset_file, #the above temporary file 
                         nhdplus_data = "download", 
                         flowline_only = FALSE, 
                         return_data = TRUE, overwrite = TRUE)
  flowline <- subset$NHDFlowline_Network
  return(flowline)
}

catchment_extract <- function(site_id){
  selectingsites <- filter(WAAIM_tidy, SITE_CD == site_id) #takes site ID and filters dataset 
  MIDLAT <- selectingsites$MIDLAT
  MIDLONG <- selectingsites$MIDLONG
  start_point <- st_sfc(st_point(c(MIDLONG, MIDLAT)), crs = 4269)
  start_comid <- discover_nhdplus_id(start_point)
  NLDI <- navigate_nldi(list(featureSource = "comid", featureID = start_comid), 
                          mode = "upstreamTributaries", 
                          distance_km = 1000)
  subset_file <- tempfile(fileext = ".gpkg")
  subset <- subset_nhdplus(comids = as.integer(NLDI$UT$nhdplus_comid), 
                         output_file = subset_file, #the above temporary file 
                         nhdplus_data = "download", 
                         flowline_only = FALSE, 
                         return_data = TRUE, overwrite = TRUE)
  catchment <- subset$CatchmentSP
  return(catchment)
}

site_name_vector <- WAAIM_tidy %>% pull(SITE_CD)

flowline_outer_frame <- data.frame(NULL)
for(cur_site_name in site_name_vector){ #took out [1:3] after site_name_vector
  cur_function_result <- flowline_extract(cur_site_name)
  head(cur_function_result)
  flowline_outer_frame <- rbind(flowline_outer_frame,cur_function_result)
}

#write.csv(outer_frame, "New Data/flowlinesall.csv")
#outer_frame %>% names() #inside of here is a thing called PATHLENGTH which is what we need to extract the distances! 

catchment_outer_frame <- data.frame(NULL)
for(cur_site_name in site_name_vector){ #took out [1:3] after site_name_vector
  cur_function_result <- flowline_extract(cur_site_name)
  head(cur_function_result)
  catchment_outer_frame <- rbind(catchment_outer_frame,cur_function_result)
}
#write.csv(outer_frame, "New Data/catchmentsall.csv")

```
#Intersecting Dam Data with flowline + catchment data above AIM sites
```{r message=FALSE, warning=FALSE}
Dam_catchment_intersect <- function(CATCHMENT, DAM_DATA){
  DAM_DATA_UTM <- st_transform(DAM_DATA, crs = 26910)
  CATCHMENT_UTM <- st_transform(CATCHMENT, crs = 26910)
  DAM_SUBSET <- DAM_DATA_UTM[lengths(
  st_intersects(DAM_DATA_UTM, CATCHMENT_UTM)) > 0,]
  return(DAM_SUBSET)
}

Dam_WAAIM_Join_function <- function(WAAIM_ROW_NUM, WAAIM_DATA, DAM_DATA){
  site_id <- WAAIM_DATA$SITE_CD[WAAIM_ROW_NUM]
  site_catchment <- catchment_extract(site_id)
  site_dams <- Dam_catchment_intersect(site_catchment, DAM_DATA)
  dams_number <- dim(site_dams)[1]
  return(dams_number)
}

result <- data.frame(matrix(nrow = 45, ncol = 2))
for(i in 1:dim(WAAIM_tidy)[1] ){
result[i,1] <-  Dam_WAAIM_Join_function(i,WAAIM_tidy,dam_complete_raw)
result[i,2] <- WAAIM_tidy$SITE_CD[i]
}
result$SITE_CD <-  result$X2
result <- result %>% 
  rename("Number_of_dams_upstream" = X1)

WAAIM_tidy_dams <- left_join(WAAIM_tidy, result)

#write.csv(WAAIM_tidy_dams, "../New Data/WAAIM_tidy_dams.csv")


Dam_complete_catchment <- Dam_catchment_intersect(catchment_outer_frame, dam_complete_raw)

#dams_upstream <- st_read("../New Data/WAAIM_tidy_dams.csv", options=c("X_POSSIBLE_NAMES=x","Y_POSSIBLE_NAMES=y"))
```




# **Exploratory Data Analysis**

To explore our data, we made simple spatial ggplots of our data sets.

# Where are the AIM sample sites and their catchments? 
```{r echo=FALSE}
ggplot(counties_sf)+
  geom_sf()+
  geom_sf(data = catchment_outer_frame, color = "blue", alpha = 0.3)+
  geom_sf(data = WAAIM_tidy, color = "green")

```
This figure shows the result of extracting the flowlines and catchments of all of the basins above each AIM site using NHDplusTools

# Where are all of dams in Washington State? 
```{r Dams in Washington State, echo=FALSE}

ggplot(counties_sf)+
  geom_sf()+
  geom_sf(data = dam_complete_raw, color = "dark red")



```
This figure is a simple plot of our Washington State Department of Ecology Inventory of Statewide Dams data.

# Catchments, Dams, and Sample Sites
```{r echo=FALSE}
ggplot(counties_sf)+
  geom_sf()+
  geom_sf(data = catchment_outer_frame, color = "blue", alpha = 0.3)+
  geom_sf(data = Dam_complete_catchment, color = "red", size = 0.5)+
  geom_sf(data = WAAIM_tidy, color = "green")


```
This is the result of intersecting all three of our data sets to map every dam that is in a watershed above an AIM site. This data is then added to the AIM data and analyzed in the next section. 

\newpage

# **Analysis**

# Correlation Check on AIM Data

We ran a correlation check to see what variables in the AIM data set are correlated with each other and also which variables are correlated with the number of dams upstream from the AIM gage site. Correlation plots can be used as a visualization to help choose what variables to include in regression models.


# Water Quality Correlation Plot
```{r correlation plots, echo=FALSE}
WAAIM_tidy_dams_merge <- read.csv("../New Data/WAAIM_tidy_dams_merge.csv")

water_quality <- WAAIM_tidy_dams_merge %>% 
  select(c(23:30, 46))

water_corr <- cor(water_quality)
corrplot(water_corr, type = "upper", diag = FALSE, tl.cex = 0.8, tl.col = "black")
```

# Biodiversity and Riparian Habitat Quality Correlation Plot
```{r echo=FALSE}
#selecting Biodiversity and Riparian Habitat Quality columns and number of dams and run correlation
#left out non-native woody veg b/c 0 for all sites
riparian_quality <- WAAIM_tidy_dams_merge %>% 
  select(c(11:14, 16:22, 46))

rip_corr <- cor(riparian_quality)
corrplot(rip_corr, type = "upper", diag = FALSE, tl.cex = 0.8, tl.col = "black")
```

# Watershed Function and Instream Habitat Quality Correlation Plot
```{r echo=FALSE}

#selecting Watershed Function and Instream Habitat Quality and number of dams and run correlation
instream_quality <- WAAIM_tidy_dams_merge %>% 
  select(c(31:46))

inst_corr <- cor(instream_quality)
corrplot(inst_corr, type = "upper", diag = FALSE, tl.cex = 0.8, tl.col = "black")
```
# Results of Correlation Plots

In the water quality plot, we saw number of dams upstream correlate positively with total nitrogen and pH. In the biodiversity and riparian habitat quality plot, we saw number of dams upstream correlate positively with non-native herbaceous vegetation. In the watershed function and in-stream habitat quality plot we saw number of dams upstream correlate negatively with percent pools in stream and average bank cover. 

# Summary and Conclusions

Washington's rivers are saturated with dams. The number of dams upstream of a sample site correlated with metrics that negatively influence water quality. However, because of the small sample size and since the AIM sites are one-off sampling occurrences, the sites provide a wealth of information but not a clear way to examine or compare data over time. Therefore it is difficult to say whether the poor water quality is caused by the dams, or if both are side-effects of highly human disturbed ecosystems. 



# Next Steps
The initial analyses we ran prepared us to further investigate the data. We plan to:
- Run an AIC to choose the most influential variables
- Run regression with the AIC determined variables
- Run a regression that includes stream order, dam type, number of dams upstream, and distance to dam with selected AIM variable as the dependent variable
- Calculate flowline distances with NHDtoolsplus
- Compare on AIM variables between sample site catchments on the basis of # of dams, dams / square mile, and flowline distance to dams





```{r correlation check, eval=FALSE, include=FALSE}
#Specify columns want to change to numeric
#Change integer columns to numeric b/c corr can only be run on numeric columns

i <- c(1, 6, 17:22, 24, 35, 40:44)  

#convert the columns indicated in i above to numeric
WAAIM_tidy[ , i] <- apply(WAAIM_tidy[ , i], 2,            # Specify own function within apply
                    function(x) as.numeric(as.character(x)))

#check that all columns we want are integers
sapply(WAAIM_tidy, class)

#drop nas so can run a corr, narrows down our data
WAAIM_tidy <- drop_na(WAAIM_tidy)

#select for numeric columns to run the corr on
WAAIM_tidy_numeric <- WAAIM_tidy %>%
  select(c(1,4:47))

#drop the geometry column b/c cor only runs on numeric columns
WAAIM_tidy_numeric <- st_set_geometry(WAAIM_tidy_numeric, NULL)

#check if geometry column removed
head(WAAIM_tidy_numeric)

#read in updated data with number of upstream dams attached. Warning is okay.
WAAIM_tidy_dams <- st_read("../New Data/WAAIM_tidy_dams.csv")

head(WAAIM_tidy_dams)

#select out the number of dams upstream column
upstream_dams <- WAAIM_tidy_dams %>% 
  select(OBJECTID, Number_of_dams_upstream)

view(upstream_dams)

#change both columns to numeric

numb <- c(1,2)  

upstream_dams[ , numb] <- apply(upstream_dams[ , numb], 2,            # Specify own function within apply
                    function(x) as.numeric(as.character(x)))

sapply(upstream_dams, class)

#merge the dfs, must be the same class to merge
WAAIM_tidy_dams_merge = WAAIM_tidy_numeric %>% left_join(upstream_dams,by="OBJECTID")
WAAIM_tidy_dams_merge

head(WAAIM_tidy_dams_merge)

# create a correlation matrix. Necessitates only complete cases. Needs to be its own data matrix. Use corr to create the matrix so all variables are the columns and rows and their correlation between each other is calculated. The dark blue diagonal down the middle is b/c when something correlated against itself is 1. Used to see how covariates are related aka correlated
WAAIMcorr <- cor(WAAIM_tidy_dams_merge)

# plot correlation matrix.
corrplot(WAAIMcorr)

# customize so that only the upper half shows since the upper and lower (compared to the diagonal mirrors itself) with the type, diag = FALSE gets the self-correlation removed, endings change the text 
corrplot(WAAIMcorr, type = "upper", diag = FALSE, tl.cex = 0.8, tl.col = "black")

#Checking correlation between variables that are categorized together in the AIM metadata.

#selecting water quality columns and number of dams and run correlation
water_quality <- WAAIM_tidy_dams_merge %>% 
  select(c(23:30, 46))

water_corr <- cor(water_quality)
corrplot(water_corr)
corrplot(water_corr, type = "upper", diag = FALSE, tl.cex = 0.8, tl.col = "black")

#selecting Biodiversity and Riparian Habitat Quality columns and number of dams and run correlation
#left out non-native woody veg b/c 0 for all sites
riparian_quality <- WAAIM_tidy_dams_merge %>% 
  select(c(11:14, 16:22, 46))

rip_corr <- cor(riparian_quality)
corrplot(rip_corr)
corrplot(rip_corr, type = "upper", diag = FALSE, tl.cex = 0.8, tl.col = "black")

#selecting Watershed Function and Instream Habitat Quality and number of dams and run correlation
instream_quality <- WAAIM_tidy_dams_merge %>% 
  select(c(31:46))

inst_corr <- cor(instream_quality)
corrplot(inst_corr)
corrplot(inst_corr, type = "upper", diag = FALSE, tl.cex = 0.8, tl.col = "black")
```





\newpage

