---
title: "Data Wrangle"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r}
getwd()
library(tidyverse)
library(lubridate)
library(sf)
theme_set(theme_classic())

```

https://hub.arcgis.com/maps/wdfw::wdfw-fish-passage-and-diversion-screening-inventory-sites/explore?location=47.280650%2C-120.845300%2C7.65 

Spatial Reference: 102100  (3857)


## Data Scrape

```{r}
dam_raw_0 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/0/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_0 <-  mutate(dam_raw_0, "Fish.Passage.Catagory" = rep("Not_a_barrier"))

dam_raw_1 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/1/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_1 <- mutate(dam_raw_1, "Fish.Passage.Catagory" = rep("Partial_Fish_Passage_Blockage"))

dam_raw_2 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/2/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_2 <- mutate(dam_raw_2, "Fish.Passage.Catagory" = rep("Total_Fish_Passage_Blockage"))

dam_raw_3 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/3/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_3 <- mutate(dam_raw_3, "Fish.Passage.Catagory" = rep("Barrier_Unknown_Percent_Passable"))

dam_raw_4 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/4/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_4 <- mutate(dam_raw_4, "Fish.Passage.Catagory" = rep("Diversion"))

dam_raw_5 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/5/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_5 <- mutate(dam_raw_5, "Fish.Passage.Catagory" = rep("Natural_Barrier_Verified"))

dam_raw_6 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/6/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_6 <- mutate(dam_raw_6, "Fish.Passage.Catagory" = rep("Natural_Barrier_Not_Field_Verified"))

dam_raw_7 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/7/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_7 <- mutate(dam_raw_7, "Fish.Passage.Catagory" = rep("On_A_Non_Fish_Bearing_Stream"))

dam_raw_8 <- st_read("https://geodataservices.wdfw.wa.gov/arcgis/rest/services/ApplicationServices/FP_Sites/MapServer/8/query?where=1%3D1&text=&objectIds=&time=&timeRelation=esriTimeRelationOverlaps&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&sqlFormat=none&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson")
dam_raw_8 <- mutate(dam_raw_8, "Fish.Passage.Catagory" = rep("Unknown"))

dam_complete_raw <- rbind(dam_raw_0, dam_raw_1, dam_raw_2,
                          dam_raw_3, dam_raw_4, dam_raw_5,
                          dam_raw_6, dam_raw_7, dam_raw_8)
rm(dam_raw_0, dam_raw_1, dam_raw_2,
    dam_raw_3, dam_raw_4, dam_raw_5,
                          dam_raw_6, dam_raw_7, dam_raw_8)

#write.csv(dam_complete_raw, "./New Data/dam_complete_raw.csv")
```

## Simple Vis of Dam Data
```{r}
ggplot(dam_complete_raw)+
  geom_sf(data = dam_complete_raw, aes(color = Fish.Passage.Catagory))

dam_complete_raw <- st_as_sf(dam_complete_raw)
class(dam_complete_raw)
```

## Read in AIM


```{r}
AIMraw <- st_read("https://gis.blm.gov/arcgis/rest/services/hydrography/BLM_Natl_AIM_AquADat/MapServer/0/query?where=1%3D1&text=&objectIds=&time=&geometry=&geometryType=esriGeometryEnvelope&inSR=&spatialRel=esriSpatialRelIntersects&distance=&units=esriSRUnit_Foot&relationParam=&outFields=*&returnGeometry=true&returnTrueCurves=false&maxAllowableOffset=&geometryPrecision=&outSR=&havingClause=&returnIdsOnly=false&returnCountOnly=false&orderByFields=&groupByFieldsForStatistics=&outStatistics=&returnZ=false&returnM=false&gdbVersion=&historicMoment=&returnDistinctValues=false&resultOffset=&resultRecordCount=&returnExtentOnly=false&datumTransformation=&parameterValues=&rangeValues=&quantizationParameters=&featureEncoding=esriDefault&f=geojson&__ncforminfo=VmJvrSyZd1Eo2ftYuiZ2Csv3FXy8H7XoE-xtqGWFdpLoXXM_QrLBbJY-BtmB02LSjkYezZWjgMjkUQTDa0n0d7ugIiEzsdKzv1tB192rp3qdK8SWS-T_Vwxqlc6whlbENi3rrg4YqK40tpTjt0dJVWg3xwroBXxhGKT92BiR7Culk8jsk8g8cxPipo1GHHZ4oF9OZUqX8NX8gOXR3zsQRZxWmOIrkcEnN_JahqrcNjjDJXXh9ysVhdPK7UcQiEvMZy-n3fIQzqq-3YFqLGCq4ffzvu0ww4Pie6NuZD2CDkf9rilS-omAyiQq1LFJ0wVAl6PywwAseMQ4f6cDJXwLDn3BcTI3WP22jrsBYXm8SzrUZs5SGhJtMOOHN1-UMdNDFrO9rqMUXOQCTyKsmhFUFerLorYSAmMVohzr7cD03LvvdSkAQXKQpHQIcFuRqac69tN3eVlyV2IfsAA8yEcpU1GJKJx4jhwaVdXX9Sv4P1ACPp9mj3yqQ5zJqgAg0aFS6Z2O4GEd2B970yfKnbeicnbkgJTSCx0BQy3gtgkxCKujAYcWeMkvV5-dYcNF7lYQRaYvX8Eod5GUIbo8N8VNPF_USB9ZCQnlesBXLVMHXSOwH_bSq49wabgeZ2ZT7aZbJ1X8M_lXzzFieFrPzSKoOK_8_Bqcfde-Hdo8Jupndsn17KsTbpSH51tbBXOC6-3NJ0oBQuZAPkZ1I3pzzqb8sTFxIpYo2ia5lSHrZ5H-gxd753QRaQJBk1xWTyWgy8mON5hZYexZrjjUH4_C_SbY0Q%3D%3D")

WAAIM <- AIMraw %>% 
  filter(PRJCT == "WA_SpokaneDO")

WAAIM_tidy <- WAAIM %>% 
  select(OBJECTID, SITE_CD, "StreamName" = STRM_NM, MIDLAT, MIDLONG, "StreamOrder" = STRM_ORDR, BTMLAT, BTMLONG, TRLAT, TRLONG, "ReachLength" = TOT_RCH_LEN, "PercentCover" = PCT_OVRHD_CVR, VEG_CMPLXTY, RPRN_VEG_CNPY_CVR, RPRN_VEG_UNDRSTRY_CVR, RPRN_VEG_GC, NON_NTVE_WDY, NTVE_WDY, NON_NTVE_HRB, NTVE_HRB, SDGE_RSH, OBSRVD_INVRT_RCHNSS, EXPCTD_INVRT_RCHNSS, MACROINVRTBRTE_CNT, "TotalN" = TTL_N, "PredictedN" = PRD_TTL_N, "TotalP" = TTL_P, "PredictedP" = PRD_TTL_P, "SpC" = SPCFC_CNDCTNCE, "PredictedSpC" = PRD_SPCFC_CNDCTNCE, pH, INSTNT_TEMP, "PercentPools" = PCT_PL, "AvgPoolDepth" = RES_PL_DEP, "NumberPools" = NM_PL, LWD_FREQ, "LWDVolume" = LWD_VL, "PercentFines2" = PCT_FN2, "PercentFines6" = PCT_FN6, D16, D84, D50, BNK_STBLTY, BNK_CVR_LWD, BNKFLL_HT, INCSN_HT, CHN_INCSN, THLWG_DEP_MN, PCT_DRY, BNKFLL_WT, WTTD_WT, "FloodWidth" = FLD_WT, "EntrenchmentRatio" = ENTRCH, SLPE, SNSTY, BVR_SGN, geometry)

rm(AIMraw)
rm(WAAIM)

getwd()
counties_sf<- st_read('./New Data/cb_2018_us_county_20m.shp') %>% 
  filter(STATEFP == 53) #Filter for just WA Counties

counties_sf_utm <- st_transform(counties_sf, crs = 26910)
dam_complete_raw_utm <- st_transform(dam_complete_raw, crs = 26910)
WAAIM_tidy_utm <- st_transform(WAAIM_tidy, crs = 26910)

```

## Simple Vis of AIM

```{r}
ggplot(data = WAAIM_tidy_utm)+
  geom_sf(data = counties_sf_utm)+
  geom_sf()

```

## Finding Dam locations that intersect the same counties as the AIM monitoring sites


```{r}
county_intersect <- counties_sf_utm[lengths(st_intersects(counties_sf_utm, WAAIM_tidy_utm)) > 0,]

ggplot(data = WAAIM_tidy_utm)+
  geom_sf(data = county_intersect)+
  geom_sf()

dam_intersect <- dam_complete_raw_utm[lengths(st_intersects(dam_complete_raw_utm, county_intersect)) > 0,]

ggplot(data = dam_intersect)+
  geom_sf(data = counties_sf_utm)+
  geom_sf()+
  geom_sf(data = WAAIM_tidy_utm, color = "red")
  

```

## Functions for flowlines and catchements from FlowWrangle

```{r}
flowline_extract <- function(site_id){
  selectingsites <- filter(WAAIM_tidy, SITE_CD == site_id) #takes site ID and filters dataset 
  MIDLAT <- selectingsites$MIDLAT
  MIDLONG <- selectingsites$MIDLONG
  start_point <- st_sfc(st_point(c(MIDLONG, MIDLAT)), crs = 4269)
  start_comid <- discover_nhdplus_id(start_point)
  NLDI <- navigate_nldi(list(featureSource = "comid", featureID = start_comid), 
                          mode = "upstreamTributaries", 
                          distance_km = 1000)
  subset_file <- tempfile(fileext = ".gpkg")
  subset <- subset_nhdplus(comids = as.integer(NLDI$UT$nhdplus_comid), 
                         output_file = subset_file, #the above temporary file 
                         nhdplus_data = "download", 
                         flowline_only = FALSE, 
                         return_data = TRUE, overwrite = TRUE)
  flowline <- subset$NHDFlowline_Network
  #catchment <- subset$CatchmentSP
  #waterbody <- subset$NHDWaterbody
  #waterlist <- list(flowline, catchment, waterbody)
  #names(waterlist) = c("flowline", "catchment", "waterbody")
  return(flowline)
}

catchment_extract <- function(site_id){
  selectingsites <- filter(WAAIM_tidy, SITE_CD == site_id) #takes site ID and filters dataset 
  MIDLAT <- selectingsites$MIDLAT
  MIDLONG <- selectingsites$MIDLONG
  start_point <- st_sfc(st_point(c(MIDLONG, MIDLAT)), crs = 4269)
  start_comid <- discover_nhdplus_id(start_point)
  NLDI <- navigate_nldi(list(featureSource = "comid", featureID = start_comid), 
                          mode = "upstreamTributaries", 
                          distance_km = 1000)
  subset_file <- tempfile(fileext = ".gpkg")
  subset <- subset_nhdplus(comids = as.integer(NLDI$UT$nhdplus_comid), 
                         output_file = subset_file, #the above temporary file 
                         nhdplus_data = "download", 
                         flowline_only = FALSE, 
                         return_data = TRUE, overwrite = TRUE)
  #flowline <- subset$NHDFlowline_Network
  catchment <- subset$CatchmentSP
  #waterbody <- subset$NHDWaterbody
  #waterlist <- list(flowline, catchment, waterbody)
  #names(waterlist) = c("flowline", "catchment", "waterbody")
  return(catchment)
}

```
testing dam intersect with extracted watersheds

```{r}
Intersect_test_1_catchment <- catchment_extract(WAAIM_tidy_utm$SITE_CD[2])
Intersect_test_1_flowline <- flowline_extract(WAAIM_tidy_utm$SITE_CD[2])
Intersect_test_1_catchment_utm <- st_transform(Intersect_test_1_catchment, crs = 26910)
Intersect_test_1_flowline_utm <- st_transform(Intersect_test_1_flowline, crs = 26910)

dam_complete_raw_utm <- st_transform(dam_complete_raw, crs = 26910)
dam_intersect_test_catch <- dam_complete_raw_utm[lengths(
  st_intersects(dam_complete_raw_utm, Intersect_test_1_catchment_utm)) > 0,]
dam_intersect_test_flow <- dam_complete_raw_utm[lengths(
  st_intersects(dam_complete_raw_utm, Intersect_test_1_flowline_utm)) > 0,]
#the dam will only intersect with the catchment data, not the flowline data. Need to trouble shoot how to add buffer to flowlines so that the dam sites are captured and can be pinned to the flowlines for calculation.

ggplot(Intersect_test_1_catchment)+
  geom_sf()+
  geom_sf(data = Intersect_test_1_flowline)+
  geom_sf(data = dam_complete_raw, aes(color = "dark red"))


```

##Writing fuction to intersect dams with catchment

```{r}
#separate function requires extract function first
Dam_catchment_intersect <- function(CATCHMENT, DAM_DATA){
  DAM_DATA_UTM <- st_transform(DAM_DATA, crs = 26910)
  CATCHMENT_UTM <- st_transform(CATCHMENT, crs = 26910)
  DAM_SUBSET <- DAM_DATA_UTM[lengths(
  st_intersects(DAM_DATA_UTM, CATCHMENT_UTM)) > 0,]
  return(DAM_SUBSET)
}

Site_5 <- catchment_extract(WAAIM_tidy$SITE_CD[5])
Site_5_Dams <- Dam_catchment_intersect(Site_5, dam_complete_raw)
ggplot(Site_5)+
  geom_sf()+
  geom_sf(data = Site_5_flow, color = "dark blue")+
  geom_sf(data = Site_5_Dams, aes(color = "dark red"))

#getting summary stats of dams and joining it to waaim_tidy
Site_5_Dams_number <- dim(Site_5_Dams)[1]
Site_5_Dams_summary <-  as.data.frame(summary(as.factor(Site_5_Dams$Fish.Passage.Catagory)))
Site_5_Dams_summary$Dam.type <- row.names(Site_5_Dams_summary)
Site_5_Dams_summary <- Site_5_Dams_summary %>% 
  pivot_wider(names_from = Dam.type, values_from = `summary(as.factor(Site_5_Dams$Fish.Passage.Catagory))`) %>% 
  mutate(SITE_CD = WAAIM_tidy$SITE_CD[5])


dam_catchment_summary <- Site_5_Dams_summary
#It works!



```

## now writing function to do the whole process start to finish

```{r}
#This function works, but only the first time it is used. Subsequent uses to aggregate data into one spreadsheet doesnt work..
Dam_WAAIM_Join_function <- function(WAAIM_ROW_NUM, WAAIM_DATA, DAM_DATA){
  site_id <- WAAIM_DATA$SITE_CD[WAAIM_ROW_NUM]
  site_catchment <- catchment_extract(site_id)
  site_dams <- Dam_catchment_intersect(site_catchment, DAM_DATA)
  #site_dams_summary <- as.data.frame(summary(as.factor(site_dams$Fish.Passage.Catagory)))  This is trying to get how many dams of each class
 # site_dams_summary$Dam.type <- row.names(site_dams_summary)
 # site_dams_summary <- site_dams_summary %>% 
    #pivot_wider(names_from = Dam.type, values_from = `summary(as.factor(site_dams$Fish.Passage.Catagory))`) %>% 
  #mutate(SITE_CD = WAAIM_DATA$SITE_CD[WAAIM_ROW_NUM])
  dams_number <- dim(site_dams)[1]
  return(dams_number)
}

#Plan is to create summary for all sites, then r bind together, then left join to waaim using site ID. I dont know how to make everything have the same number of rows if there are missing values...
dam_catchment_summary_1 <- Dam_WAAIM_Join_function(1,WAAIM_tidy,dam_complete_raw)
dam_catchment_summary_2 <- Dam_WAAIM_Join_function(2,WAAIM_tidy,dam_complete_raw)
dam_catchment_summary_3 <- Dam_WAAIM_Join_function(3,WAAIM_tidy,dam_complete_raw)
dam_catchment_summary_4 <- Dam_WAAIM_Join_function(4,WAAIM_tidy,dam_complete_raw)
dam_catchment_summary_5 <- Dam_WAAIM_Join_function(5,WAAIM_tidy,dam_complete_raw)


result <- data.frame(matrix(nrow = 45, ncol = 2))
for(i in 1:dim(WAAIM_tidy)[1] ){
result[i,1] <-  Dam_WAAIM_Join_function(i,WAAIM_tidy,dam_complete_raw)
result[i,2] <- WAAIM_tidy$SITE_CD[i]
}
#THE FOR LOOP WORKED AND I FEEL SO POWERFUL

result$SITE_CD <-  result$X2
result <- result %>% 
  rename("Number_of_dams_upstream" = X1)

WAAIM_tidy_dams <- left_join(WAAIM_tidy, result)

write.csv(WAAIM_tidy_dams, "./New Data/WAAIM_tidy_dams.csv")

dams_upstream <- st_read("./New Data/WAAIM_tidy_dams.csv", options=c("X_POSSIBLE_NAMES=x","Y_POSSIBLE_NAMES=y"))
```




