---
title: "Flow Lines"
author: "Jackie Van Der Hout"
date: "4/5/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
getwd()
library(tidyverse)
# install.packages("nhdplusTools") #uncomment and run!
library(nhdplusTools)
library(sf)
library(tmap)
tmap_mode("view") #makes map plots interactive
library(dataRetrieval)
library(maps)

theme_set(theme_classic())
options(scipen = 5) 
```

example: getting the flowlines for one AIM site
```{r}
# Extract latitude and longitude for each site

WAAIMcoords1 <- WAAIM_tidy %>% 
  filter(SITE_CD == "SW-SS-19199") %>% 
  select(SITE_CD, MIDLAT, MIDLONG)

# Define the sample site as the starting point
start_point <- st_sfc(st_point(c(WAAIMcoords1$MIDLONG, WAAIMcoords1$MIDLAT)), 
                      crs = 4269) # NAD83

start_comid <- discover_nhdplus_id(start_point)

NLDI <- navigate_nldi(list(featureSource = "comid", featureID = start_comid), 
                          mode = "upstreamTributaries", 
                          distance_km = 1000)

# Extract watershed and flowpath information
subset_file <- tempfile(fileext = ".gpkg")
subset <- subset_nhdplus(comids = as.integer(NLDI$UT$nhdplus_comid), 
                         output_file = subset_file, #the above temporary file 
                         nhdplus_data = "download", 
                         flowline_only = TRUE, 
                         return_data = TRUE, overwrite = TRUE)

# Create data frame for flowline
flowline <- subset$NHDFlowline_Network
class(flowline) #a spatial object

#plotting successes
states <- st_as_sf(maps::map(database = "state", plot = FALSE, fill = TRUE, col = "white"))
wa <- filter(states, ID == "washington")

ggplot(wa) +
  geom_sf(fill = "white") +
  geom_sf(data = flowline, aes(color = streamorde)) +
  labs(color = "Stream Order") +
  theme(legend.position = "top")
#HAHAHAHAHA

ggplot(flowline) +
  geom_sf(fill = "white", color = "gray", lwd = 0.5) +
  geom_sf(data = flowline, aes(color = streamorde)) +
  labs(color = "Stream Order") +
  theme(legend.position = "top")
#WE HAVE A PLOT!!!

#plotting failures
#tm_shape(flowline)+tm_lines() getting an error path too long

#tm_shape(UT_flowlines) + tm_lines() +
  #tm_shape(UT_nwissite) +tm_dots(col="red") 

#plot_nhdplus(WAAIMcoords1$MIDLONG, WAAIMcoords1$MIDLAT, streamorder = 3)
```


writing the function
**NOTE** 
this function has been tested and works!!!!
USE THESE FUNCTIONS!!!
```{r}
  
flowline_extract <- function(site_id){
  selectingsites <- filter(WAAIM_tidy, SITE_CD == site_id) #takes site ID and filters dataset 
  MIDLAT <- selectingsites$MIDLAT
  MIDLONG <- selectingsites$MIDLONG
  start_point <- st_sfc(st_point(c(MIDLONG, MIDLAT)), crs = 4269)
  start_comid <- discover_nhdplus_id(start_point)
  NLDI <- navigate_nldi(list(featureSource = "comid", featureID = start_comid), 
                          mode = "upstreamTributaries", 
                          distance_km = 1000)
  subset_file <- tempfile(fileext = ".gpkg")
  subset <- subset_nhdplus(comids = as.integer(NLDI$UT$nhdplus_comid), 
                         output_file = subset_file, #the above temporary file 
                         nhdplus_data = "download", 
                         flowline_only = FALSE, 
                         return_data = TRUE, overwrite = TRUE)
  flowline <- subset$NHDFlowline_Network
  #catchment <- subset$CatchmentSP
  #waterbody <- subset$NHDWaterbody
  #waterlist <- list(flowline, catchment, waterbody)
  #names(waterlist) = c("flowline", "catchment", "waterbody")
  return(flowline)
}

catchment_extract <- function(site_id){
  selectingsites <- filter(WAAIM_tidy, SITE_CD == site_id) #takes site ID and filters dataset 
  MIDLAT <- selectingsites$MIDLAT
  MIDLONG <- selectingsites$MIDLONG
  start_point <- st_sfc(st_point(c(MIDLONG, MIDLAT)), crs = 4269)
  start_comid <- discover_nhdplus_id(start_point)
  NLDI <- navigate_nldi(list(featureSource = "comid", featureID = start_comid), 
                          mode = "upstreamTributaries", 
                          distance_km = 1000)
  subset_file <- tempfile(fileext = ".gpkg")
  subset <- subset_nhdplus(comids = as.integer(NLDI$UT$nhdplus_comid), 
                         output_file = subset_file, #the above temporary file 
                         nhdplus_data = "download", 
                         flowline_only = FALSE, 
                         return_data = TRUE, overwrite = TRUE)
  #flowline <- subset$NHDFlowline_Network
  catchment <- subset$CatchmentSP
  #waterbody <- subset$NHDWaterbody
  #waterlist <- list(flowline, catchment, waterbody)
  #names(waterlist) = c("flowline", "catchment", "waterbody")
  return(catchments)
}

testtest <- flowline_extract("SW-SS-19199")


```


automating the extraction of the flowlines and catchment for the 45 sites? 
```{r}
#workshopping this 
site_name_vector <- WAAIM_tidy %>% pull(SITE_CD)

flowline_outer_frame <- data.frame(NULL)
for(cur_site_name in site_name_vector){ #took out [1:3] after site_name_vector
  cur_function_result <- flowline_extract(cur_site_name)
  head(cur_function_result)
  flowline_outer_frame <- rbind(outer_frame,cur_function_result)
}

write.csv(outer_frame, "New Data/flowlinesall.csv")
outer_frame %>% names() #inside of here is a thing called PATHLENGTH which is what we need to extract the distances! 

catchment_outer_frame <- data.frame(NULL)
for(cur_site_name in site_name_vector){ #took out [1:3] after site_name_vector
  cur_function_result <- flowline_extract(cur_site_name)
  head(cur_function_result)
  catchment_outer_frame <- rbind(outer_frame,cur_function_result)
}
write.csv(outer_frame, "New Data/catchmentsall.csv")

```

some viz
```{r}
tmap_mode("view")
tm_shape(outer_frame) + tm_lines() +
  tm_shape(WAAIM_tidy) + tm_dots() 
  #tm_shape(dam_complete_raw)+ tm_dots(col="red")

#need to clip the dam data to only be present for the sample site catchments 
```


Intersecting catchment with dam data
```{r}
#Now intersecting catchment areas with dam data
dam_complete_raw_crs <-  st_transform(dam_complete_raw, crs = 4269)
TestRunintersect <- dam_complete_raw_crs[lengths(st_intersects(dam_complete_raw_crs, TestRunFlowline)) > 0,]

ggplot(TestRunFlowline) +
    geom_sf(fill = "white", color = "gray", lwd = 0.5, alpha = 0.5) +
    geom_sf(data = TestRunintersect, aes(color = Fish.Passage.Catagory))+
    #geom_sf(data = TestRunCatchment) +
    labs(color = "Stream Order") 

```

extracting lengths between dams and AIM on flowline
Part One 1:
---
Dam ID | COM_ID | Dam_Path_Length | AIM_ID
---
Part 2:
AIM_SI | COM_ID | Dam_Path_Length
---
THEN joing the two on ??
```{r}
#get the com_ID for a lat long then attach it to anything in the NHD
#you really want the path length, which is the distance to the outlet 
#get the aim site pathway, and take the difference between the two 
#that is the distance 
mutate(DistanceToDam = Dam_Path_Length - AIM_Path_Length)

st_nearest_feature()
#can be used on dam sites to return the COM_ID that is nearest to it 

#how to attach to COM_ID

#you can use those catchment polygons to subset at dataset 
#polygons can be joined to represent a whole aim site 

```


Indexing flowlines 
**NOTE** 
This code is from: https://usgs-r.github.io/nhdplusTools/articles/indexing.html
And has not been edited for our data
```{r}

# \donttest{

source(system.file("extdata", "sample_flines.R", package = "nhdplusTools"))

point <- sf::st_sfc(sf::st_point(c(-76.87479, 39.48233)),
                    crs = 4326)

get_flowline_index(sample_flines, point)

point <- sf::st_transform(point, 5070)

get_flowline_index(sample_flines, point,
                   search_radius = units::set_units(200, "m"))

get_flowline_index("download_nhdplusv2", point)

get_flowline_index(sample_flines, point, precision = 30)

get_flowline_index(sample_flines,
                   sf::st_sfc(list(sf::st_point(c(-76.86934, 39.49328)),
                                   sf::st_point(c(-76.91711, 39.40884)),
                                   sf::st_point(c(-76.88081, 39.36354))),
                              crs = 4326),
                   search_radius = units::set_units(0.2, "degrees"),
                   max_matches = 10)

# }


indexes <- get_flowline_index(sf::st_transform(flowlines, 5070), # albers
                              sf::st_transform(sf::st_geometry(gages), 5070), 
                              search_radius = units::set_units(200, "meters"), 
                              max_matches = 1)

indexes <- left_join(sf::st_sf(id = c(1:nrow(gages)), 
                               geom = sf::st_geometry(gages)), 
                     indexes, by = "id")

plot(sf::st_geometry(sf::st_zm(flowlines)))
plot(sf::st_geometry(indexes), add = TRUE)


```


Initial notes onCalculating River Distance Options
```{r}
#river dist package https://github.com/mbtyers/riverdist

#alternate river distance calculation method https://fsdias.github.io/flow_distance.md/
#river mile calculator outside of R https://waterdata.usgs.gov/wa/nwis/current?type=rivermi

#bin sites into upstream or downstream of dam
#add column with mutate function correlating above or below dam
#another column for distance 
#to calculate that distance, it needs to be calculated along the stream 
#to calculate that use the riverdist package https://github.com/mbtyers/riverdist
#within that there is the river distance function
#I will need to grab the shapefiles (background map) 
#NHD+ tools will have the background map with the streams https://usgs-r.github.io/nhdplusTools/
#bin directly downstream and distance downstream 
#make sure that all of the files are in the same projection before doing the analysis #projection of the basemap is what the other two files should be in 
```

**WARNING**
*unnecessarily complicated code and one giant download*
*do not attempt unless you have an hour for your machine to freak out*
ndhplus & intersectr attempt ONE
https://github.com/USGS-R/intersectr
```{r}
library(nhdplusTools)
library(sf)
site <- list(featureSource = "nwissite",
               featureID = "USGS-10128500")
line <- navigate_nldi(site, "UT", "")
site <- navigate_nldi(site, "UT", "nwissite")


                      nhdp <- subset_nhdplus(ut$nhdplus_comid,
                       "nhdp_subset.gpkg",
                      "download")
  
nhdplusTools::get_flowline_index(
  points = sites, 
  flines = flowline,
  searchradius = 100, #units of flowlines
  precision = 10) #max node spacing

#spatial intersection
#could be used to intersect dam and sample locations to flow lines? 
#install.packages("devtools")
#library(devtools)
#be forewarned -- HUGE DOWNLOAD!!!!
#devtools::install_github("USGS-R/intersectr")
#library(intersectr)
cells <- create_cell_geometry(x_coords, y_coords, nc_projections, catchment)
poly <- sf::st_as_sf(dplyr::select(catchment, ID = featureid))
weights <- calculate_area_intersection_weights(cells, poly)
dap_uri <- "#metadatalink"
execute_intersection(nc_file = dap_uri,
                     variable name = #"precipitation_amount", 
                intersection_weights = weights,
                cell_geometry = data_source_cells,
                x_var = x_coord, y_var = y_coord, t_var = t_coord, 
                start_datetime = #"",
                end_datetime = #"")

```

WARNING
**STALE CODE BELOW**
*functional but unnecessary code, has been improved above*

```{r}
#unsure if flowline, catchment, waterbody need a command to create custom name for each different site

####does this function allow us to track which site is which for futher analysis? 
extract_flowlines_plot_WAAIM_tidy <- function(ROWNUM){
  MIDLAT <- WAAIM_tidy$MIDLAT[ROWNUM]
  MIDLONG <- WAAIM_tidy$MIDLONG[ROWNUM]
  start_point <- st_sfc(st_point(c(MIDLONG, MIDLAT)), crs = 4269)
  start_comid <- discover_nhdplus_id(start_point)
  NLDI <- navigate_nldi(list(featureSource = "comid", featureID = start_comid), 
                          mode = "upstreamTributaries", 
                          distance_km = 1000)
  subset_file <- tempfile(fileext = ".gpkg")
  subset <- subset_nhdplus(comids = as.integer(NLDI$UT$nhdplus_comid), 
                         output_file = subset_file, #the above temporary file 
                         nhdplus_data = "download", 
                         flowline_only = FALSE, 
                         return_data = TRUE, overwrite = TRUE)
  flowline <- subset$NHDFlowline_Network
  catchment <- subset$CatchmentSP
  waterbody <- subset$NHDWaterbody
  cur_plot <-  ggplot(flowline) +
    geom_sf(fill = "white", color = "gray", lwd = 0.5) +
    geom_sf(data = flowline, aes(color = streamorde)) +
    labs(color = "Stream Order") +
    theme(legend.position = "top")
  print(cur_plot)
  return(line)
}

#test of function
extract_flowlines_plot(1)
extract_flowlines(2)
#I think it works!

```

Functions to save shapefiles of catchement and flowlines

```{r}
#File to save the shapefile of flowlines with

extract_flowlines_file_WAAIM_tidy <- function(ROWNUM){
  MIDLAT <- WAAIM_tidy$MIDLAT[ROWNUM]
  MIDLONG <- WAAIM_tidy$MIDLONG[ROWNUM]
  start_point <- st_sfc(st_point(c(MIDLONG, MIDLAT)), crs = 4269)
  start_comid <- discover_nhdplus_id(start_point)
  NLDI <- navigate_nldi(list(featureSource = "comid", featureID = start_comid), 
                          mode = "upstreamTributaries", 
                          distance_km = 1000)
  subset_file <- tempfile(fileext = ".gpkg")
  subset <- subset_nhdplus(comids = as.integer(NLDI$UT$nhdplus_comid), 
                         output_file = subset_file, #the above temporary file 
                         nhdplus_data = "download", 
                         flowline_only = FALSE, 
                         return_data = TRUE, overwrite = TRUE)
  flowline <- subset$NHDFlowline_Network
  return(flowline)
}

#test
Site_1 <-  extract_flowlines_file_WAAIM_tidy(1)
Site_2 <-  extract_flowlines_file_WAAIM_tidy(2)


#attmepting to extract all at once, getting an error
therows <- rep(1:45)

extract_flowlines_file_WAAIM_tidy(therows) %>% 
  bind_rows()

#File to save the shapefile of catchments with
extract_catchments_file_WAAIM_tidy <- function(ROWNUM){
  MIDLAT <- WAAIM_tidy$MIDLAT[ROWNUM]
  MIDLONG <- WAAIM_tidy$MIDLONG[ROWNUM]
  start_point <- st_sfc(st_point(c(MIDLONG, MIDLAT)), crs = 4269)
  start_comid <- discover_nhdplus_id(start_point)
  NLDI <- navigate_nldi(list(featureSource = "comid", featureID = start_comid), 
                          mode = "upstreamTributaries", 
                          distance_km = 1000)
  subset_file <- tempfile(fileext = ".gpkg")
  subset <- subset_nhdplus(comids = as.integer(NLDI$UT$nhdplus_comid), 
                         output_file = subset_file, #the above temporary file 
                         nhdplus_data = "download", 
                         flowline_only = FALSE, 
                         return_data = TRUE, overwrite = TRUE)
  catchment <- subset$CatchmentSP
  return(catchment)
}

#Test
Site_1_catchment <- extract_catchments_file_WAAIM_tidy(1)
Site_4_catchment <- extract_catchments_file_WAAIM_tidy(4)
Site_4_flow <- extract_flowlines_file_WAAIM_tidy(4)

#it works!!!
ggplot(Site_4_catchment) +
    geom_sf(fill = "white", color = "gray", lwd = 0.5) +
    geom_sf(data = Site_4_flow, aes(color = streamorde)) +
    labs(color = "Stream Order") +
    theme(legend.position = "top")
```

