---
title: "NHD data expedition day"
author: "Nick Bruns"
date: "2/18/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

Consult [these docs from this Download link](https://www.horizon-systems.com/NHDPlusData/NHDPlusV21/Documentation/NHDPlusV2_User_Guide.pdf) to understand NHD column values, their units etc.
This is long, so just control-f on the information you need.

[This map has all USGS gages](https://maps.waterdata.usgs.gov/mapper/index.html) 

[This tool is nice for visualizing HUCs](http://waterqualityexplorer.rc.duke.edu:3838/explorer/) 



```{r setup, include=FALSE}
# install.packages("nhdplusTools") #uncomment and run!
library(nhdplusTools)
library(tidyverse)
library(sf)
library(tmap)
tmap_mode("view") #makes map plots interactive
```

---


```{r target_hucs}
#some gage ID's
New_Hope_Creek_gage_id <- "02097314"
Eno_gage_id <- "02085070"
Yampa_gage_id <- "09244490"
Yampa_gage_id_upstream <- "09239500"
Weber_river_id <- "10128500" #UTah
Salmon_river_ADK_id <- "04249200" #Aderondacks (good clean Hydrograph!)
Salmon_river_ID_head_waters<- "13310700"
Yellowstone_river_id <- "06214500" # at Billings
Flint_river <- "02349605" #Georgia

```

problem in this chunk with plotting, cur_line is list not sf?
I ended up extracting the two items from the lists of both and trying to map that instead
and did get a map!
how do i make this work with lat / long instead of usgs gages?
```{r code_demonstration}
#First, we'll check how many machines work. So first, run this whole block, then we'll work through line by line.
  cur_gage_id <- Eno_gage_id
  cur_id_string<- paste0("USGS-",cur_gage_id) #just add prefix to string
  cur_site <- list(featureSource = "nwissite",  #prepping below web quirie below, this just indicates the kind of data we have
               featureID = cur_id_string) 
  cur_line <- navigate_nldi(cur_site, "UT", "") #returns all upstream flowlines
  cur_site_sf <- navigate_nldi(cur_site, "UT", "nwissite") #returns all upstream USGS gages

  class(cur_line) #list
#st_as_sf(cur_line, fill = TRUE, group = TRUE)
  
cur_line_origin <- cur_line$origin
UT_flowlines <- cur_line$UT_flowlines
cur_site_sf_origin <- cur_site_sf$origin
UT_nwissite <- cur_site_sf$UT_nwissite
  
#now map flowlines and gages
tm_shape(UT_flowlines) + tm_lines() +
  tm_shape(UT_nwissite) +tm_dots(col="red") 
```


for this section, should i be using cur_line or UT_flowlines in subset_nhdplus()? 
I went ahead and made the substitution
```{r download_NHD_info}
#Flowlines above are just the ID's . Download everything!
save_file_name <- "Eno_River_NHD_subset.gpkg" #gkpg is an open source geospatial format

class(UT_flowlines$nhdplus_comid)
UT_flowlines$nhdplus_comid <- as.numeric(UT_flowlines$nhdplus_comid)
class(UT_flowlines$nhdplus_comid)

subset_nhdplus(UT_flowlines$nhdplus_comid, 
                       save_file_name, 
                       "download")

cur_nhd <- st_read(save_file_name)
glimpse(cur_nhd)
```

```{r make_nhd_plot}
cur_density <- sum(cur_nhd$lengthkm)/max(cur_nhd$totdasqkm) #network density metric
print(cur_density)

#plot elevation
cur_nhd %>% 
  # filter(pathlength>2100) %>% #Necessary for Yampa river
  ggplot() + 
  geom_line(aes(x=pathlength,y=minelevsmo,col=as.factor(streamorde))) +
   xlab("km to outlet") +
  ylab("elevation (cm)") +
  theme_light() + scale_color_viridis_d("stream order")
```

I want to customize this function to work with a list of AIM and dam sites! How do I do this if it isn't connected to USGS gages??
Also, is it the sample sites or the dam sites that I want to input here? 
```{r wrap_all_above_code_into_functions}
#great. That's it. Now we'll wrap all in functions for easier  use.
plot_basin <- function(gage_id_string){
  id_string=paste0("USGS-",gage_id_string) 
  site <- list(featureSource = "nwissite", 
                featureID = id_string)
  print("getting site") 
  discover_nldi_navigation(site)
  print("getting flowline")  
  line <- navigate_nldi(site, "UT", "") #returns all upstream comids
  print("getting usptream sites")
  site_sf <- navigate_nldi(site, "UT", "nwissite") #returns all upstream gages!
  cur_plot <- tm_shape(line) + tm_lines() +
    tm_shape(site_sf) +tm_dots(col="red") 
  print(cur_plot)
  return(line)
}

download_slim_nhd <- function(flow_line_with_comid,save_string){
  subset_nhdplus(flow_line_with_comid$nhdplus_comid, 
                       save_string, 
                       "download")
}

plot_elevation <- function(cur_nhd,river_name){
  cur_nhd %>% 
  ggplot() + 
  geom_line(aes(x=pathlength,y=minelevsmo,col=as.factor(streamorde))) +
   xlab("km to outlet") +
  ylab("elevation (cm)") +
  theme_dark() + scale_color_viridis_d("stream order") +
    ggtitle(river_name)
}

#Geomorphic Instantenous Unit Hydrograph
plot_GIUH <- function(cur_nhd){
  ggplot(cur_nhd) +
    geom_point(aes(y=totdasqkm,x=pathlength,col=as.character(streamorde))) + scale_color_viridis_d("stream order") +
    theme_dark() 
}

compute_network_density<- function(cur_nhd){ 
  sum(cur_nhd$lengthkm)/max(cur_nhd$totdasqkm)
}

plot_stream_order <- function(cur_nhd,matching_pal=T) {
  if(matching_pal){
    cur_nhd %>% mutate(streamorde=as.factor(streamorde)) %>%  
      tm_shape() + tm_lines(col="streamorde",lwd = 3,palette = "viridis")
  } else{
     cur_nhd %>% mutate(streamorde=as.factor(streamorde)) %>%  
      tm_shape() + tm_lines(col="streamorde",lwd = 3)
  }
  
}
```
```{r show_function_use_example}
#setup: to try a different basin, 
# insert new code chunk (cmd + option + i)
# paste this code chunk
# modify this top part with a new name and USGS gage id.
cur_river_name <- "New_Hope_Creek"
cur_gage_id <- New_Hope_Creek_gage_id #or directly put a quoted gage id here
##could we also put lat long here instead??
print(cur_gage_id)

#run code
cur_file_name <- paste0(cur_river_name,"_NHD_subset.gpkg")
cur_flow_line <- plot_basin(cur_gage_id)
download_slim_nhd(cur_flow_line,cur_file_name) #will fail if the file already exists
##error: could not find function "discover_nldi_navigation"

cur_nhd <- st_read(cur_file_name) #load that NHD
compute_network_density(cur_nhd)
plot_elevation(cur_nhd,cur_river_name)
plot_stream_order(cur_nhd )
plot_GIUH(cur_nhd )
```









